\documentclass[11pt]{article}
\usepackage[margin=.8in,letterpaper]{geometry}

\usepackage{times}
\usepackage{enumitem}
\usepackage{titlesec}
\renewcommand{\familydefault}{\sfdefault}

\titleformat*{\section}{\large\bfseries}

\title{\vspace{-0.2in}\textbf{Advanced Placement Physics}}
\author{Instructor: Dr.\ Timothy Leung (\texttt{tleung@olympiadsmail.ca})}
%\date{Fall 2018}
\date{Summer 2019}

\begin{document}
\maketitle

\subsection{Class Time}
%\begin{itemize}[itemsep=.1em,leftmargin=12pt]
%\item Saturdays 4:20pm--6:50pm (Starts November 3, 2018)
Saturdays \& Sundays 7:00--10:00pm (July and August)\\
Class time in September TBA
%\end{itemize}

\subsection{Course Material}
\begin{itemize}[itemsep=0pt,leftmargin=12pt]
\item No textbook required
\item Please check the school website for lecture slides, homework, and other
  resources.
\item Students are expected to bring the following to each class:
  \begin{itemize}[noitemsep,topsep=0pt]
  \item A pen/pencil for note-taking
  \item Paper/notebook/binder
  \item A scientific calculator for working in-class example problems
  \end{itemize}
\end{itemize}

\subsection{Classroom Expectations}
Students are expected to:
\begin{itemize}[itemsep=0pt,leftmargin=12pt]
\item Be in your seat and ready to learn and participate during class.
\item Stay on task without disturbing or distracting others.
\item Raise your hand if you have any questions or comments and wait to be
  called. Don't wait too long before you ask a question.
\item If you need to leave the class early, your parent needs to pick you up at
  the classrom door.
\item Be respectful for yourself, others, and the facilities; act in
  a responsible manner in everything you do.
\end{itemize}

\subsection{Homework Expectation}

\begin{itemize}[itemsep=0pt,leftmargin=12pt]
\item Homework is assigned approximately every \emph{two} weeks, depending on
  the material being taught.
\item Not regularly completing the homework prompts a phone call to communicate
  with the parents in order to help you better manage your time and achieve
  your goal.
\item Late homework is accepted. 
\item For free-response questions:
  \begin{itemize}[noitemsep]
  \item Show \emph{all} work by providing complete and organized steps. Answer
    the questions as if the reader is learning the concept from you, not as if
    s/he already understands it.
  \item If a question requires you to \emph{explain}, please do so using
    complete sentences with supporting detail.
  \item Proper math format must be used, e.g.\ proper use of ``='' sign, units,
    etc.
  \item Circle or box all your final answers.
  \end{itemize}
\item I will take up some of the more difficult questions during class.
  However, this does \emph{not} mean you don't need to do your homework at
  home. Always do your best.
\item Please keep all returned homework assignmentsin your binder until the end
  of the term.
\item The homework is marked with ``P'' for pass and ``I'' for incomplete. If
  you do get an ``I'', I will let you know how to fix it so you can re-submit
  to pass the homework. 
\end{itemize}


\subsection*{Pre-requisites}
  \begin{itemize}[itemsep=0pt,leftmargin=12pt]
  \item\textbf{Physics 11 and 12:} Student will need to be competent in all the
    topics covered in the high-school level courses. Many topics from Phyiscs
    11 and 12 are covered more in-depth in this course.
    covered
  \item\textbf{Calculus:} The two ``C'' exams are calculus based, and students
    are required to perform basic differentiation and integration.
  \item\textbf{Vectors:} Students need to have basic understanding of vector
    operations, including addition and subtraction, as well as dot products and
    cross products.
  \end{itemize}



  \subsection*{Course Outline}
%The course covers all topics in AP Physics 1, 2, and C exams.
%\begin{center}
%  \begin{tabular}{|c|l|c|}
%    \textbf{Class} & \textbf{Topic} & Exams \\ \hline
%    1 & Kinematics using calculus & 1 \& C (Mechanics)\\
%    2 & Dynamics & 1 \& C (Mechanics)\\
%    3,4 & Momentum, impulse and energy &  1 \& C (Mechanics)\\
%    4,5  & Center of mass & 1 \& C (Mechanics)\\
%    5,6 & General circular motion &  1 \& C (Mechanics)\\
%    7,8 & Universal gravitation, Kepler's laws and planetary motion &
%  1 \& C (Mechanics)\\
%  9 & Practice Test \# 1 (AP Physics C: Mechanics) & \\ \hline
%  10 & Electrostatics & 2 \& C (Electricity \& Magnetism)\\
%  11 & Gauss's law \& capacitors & 2 \& C (Electricity \& Magnetism)\\
%  12,13 & Circuits analysis with resistors and capicators &
%  2 \& C (Electricity \& Magnetism)\\
%  14,15 & Magnetism & 2 \& C (Electricity \& Magnetism)\\
%  16 & Maxwell's laws of electromagnetism & 2 \& C (Electricity \& Magnetism)\\
%  17 & Practice Test \# 2 (AP Physics C: Electricity and Magnetism) &\\
%  \hline
%  18
%  \end{tabular}
%\end{center}
\begin{enumerate}[itemsep=0pt,leftmargin=15pt]
\item Topics in \emph{AP Physics C: Mechanics}
  \begin{enumerate}[itemsep=0pt,leftmargin=18pt]
  \item Kinematics
  \item Dynamics
  \item Momentum, impulse and energy
  \item Center of mass
  \item General circular motion and angular momentum
  \item Simple harmonic motion--general equation of oscillatory systems,
    pendulums and spring-mass systems
  \item Universal gravitation and planetary motion
  \item\textbf{Practice AP Physics C: Mechanics exam}
  \end{enumerate}
\item Topics in \emph{AP Physics C: Electricity and Magnetism} (``E\&M'')
  \begin{enumerate}[itemsep=0pt,leftmargin=18pt]
  \item Electrostatics
  \item Gauss's law
  \item Capacitance
  \item Magnetism
  \item Inductance
  \item Circuit analysis (RC, RL, LC and RLC circuits)
  \item Maxwell's equations and electromagnetic wave
  \item\textbf{Practice AP Physics C: E\&M exam}
  \end{enumerate}
\item Additional topics in \emph{AP Physics 1} and \emph{AP Physics 2}
  \begin{enumerate}[itemsep=0pt,leftmargin=18pt]
  \item Fluid dynamics
  \item Thermal physics
  \item Mechanical waves
  \item Light and optics
  \item Special relativity
  \item Quantum mechanics
  \item\textbf{Practice AP Physics 2 Exam}
  \end{enumerate}
\end{enumerate}
\end{document}


