\documentclass[11pt]{article}
\usepackage{bm}
\usepackage[margin=1.7cm,top=2.5cm,bottom=2.5cm,letterpaper]{geometry}
\usepackage{enumitem}
\usepackage{tikz,graphicx,wrapfig}
\usepackage{amsmath}
%\usepackage{pgfplots}
\usepackage{times}
\usepackage[document]{ragged2e}
\usepackage[none]{hyphenat}
\usepackage{siunitx}
\usepackage{multicol}
\usepackage{fancyhdr}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\iii}{\bm{\hat{\imath}}}
\newcommand{\jjj}{\bm{\hat{\jmath}}}

\setlength{\parskip}{15pt}

\sisetup{
  %  mode=text,
  detect-all,
  %number-math-rm=\mathnormal,
  %  text-sf=\sffamily,
  %  text-rm=\sffamily,
  inter-unit-product =\ensuremath{\cdot{}},
  per-mode=symbol,
  group-separator={,},
}

\pagestyle{fancy}
  \chead{}
  \lhead{Olympiads School}
  \rhead{Solutions to Free Response Questions}
  \lfoot{Advanced Placement Physics}
  \cfoot{-\thepage-}
  \rfoot{}


\begin{document}

\fbox{
  \begin{minipage}{.97\linewidth}
    \textbf{Free Response Question 4:} A steel ball is dropped from a point with
    $(x,y)$ coordinate of $(\SI{8}{\metre},\SI{16}{\metre})$. At the same time,
    another ball is launched from the origin with a speed of
    \SI{20}{\metre\per\second} at an angle of \ang{30}.
    \begin{enumerate}[noitemsep,topsep=0pt]
    \item Find the minimum distance of separation occur of the two balls.
    \item At what time does this separation occur?
    \item Give the coordinates of the two balls for the minimum separation.
    \end{enumerate}
  \end{minipage}
}

There are two ways to solve the problem. The first method \emph{looks} easy on
first glance, but will require a lot of calculus. The second method, on the
other hand, is a straightforward geometry problem that requires a bit of
ingenuity.

\textbf{\underline{Method 1 (not recommended)}}

The most straightforward approach is to express the distance between the steel
balls as a function of time, and then take the derivative with respect to time
to find out when it occurs $t$ and minimum value of $d$.

Let's call the steel ball being dropped $A$, and the ball that is launched
$B$. Their respective position in the coordinate system are expressed as
functions of time using kinematic equations\footnote{For the $x$ direction,
  $x=v_xt$ and for the $y$ directiom, $y=y_0+v_{y0}t-\frac12gt^2$.}, and using
$g=\SI{10}{\metre\per\second^2}$ for both cases\footnote{which is acceptable
  for all AP exams} for simplicity.
\begin{align}
  \mb{x}_A &= 8\iii + (16-5t^2)\jjj \\
  \mb{x}_B &=20\cos\ang{30}t\iii+(20\sin\ang{30}t-5t^2)\jjj
\end{align}
The ``displacement'' vector between $A$ and $B$ can be expressed as:
\begin{equation}
  \Delta\mb{x}
  =\mb{x}_A-\mb{x}_B
  =(8-20\cos\ang{30}t)\iii + (16-20\sin\ang{30}t)\jjj
  \label{no-g}
\end{equation}
Not surprisingly, the gravitational acceleration term $\frac12gt^2=5t^2$ term
cancels, because both $A$ and $B$ are free-falling objects with the same
downward acceleration. The square of the \emph{distance} between $A$ and $B$ are
expressed as:
\begin{equation}
  d^2 =(8-20\cos\ang{30}t)^2 + (16-20\sin\ang{30}t)^2
\end{equation}
What we will need to do now is to take the time derivative of $d^2$ with
respect to time, and to find $d$ and $t$. This process is laborious and
tedius (and prone to error for someone uncomfortable with using chain rule)
and therefore generally not recommended. We will instead try a completely
different approach.

\textbf{\underline{Method 2}}

However, we have already noted that in Equation~\ref{no-g}, acceleration due to
gravity terms cancel, which means that the minimum separation distance $d$ does
not depend on $g$! We instead consider an observer who is falling alongside
steel ball $A$ (which is equivalent to effectively treating $g=0$.) In this
case, the observer sees that $A$ remains stationary while $B$ travels in a
straight line instead of the parabolic path of a projectile. The observer's
point of view is shown in Fig.~\ref{falling}. The minimum distance of
separation occurs at $C$ in this frame of reference, with a value of $d$.
\begin{figure}[ht]
  \begin{center}
    \begin{tikzpicture}[scale=.4]
      \draw[->](0,0)--(17,0) node[pos=1,right]{$x$};
      \draw[->](0,0)--(0,17) node[pos=1,above]{$y$};
      \fill(8,16) circle(.18) node[above]{$A$};
      \fill(8,4.62) circle(.1) node[above left]{$E$};
      \fill(8,0) circle(.1) node[below]{$D$};
      \draw[dashed](8,16)--(8,0);
      \draw[dashed](12.9,7.46)--(12.9,0) node[below]{$C'$};
      \draw[->](3,0) arc(0:30:3) node[midway,right]{\ang{30}};
      \draw[->](8,13) arc(270:300:3) node[midway,below]{\ang{30}};
      \begin{scope}[rotate=30]
        \draw[very thick](0,0)--(18,0);
        \fill(18,0) circle(.18) node[right]{$B$};
      \end{scope}
      \draw[dashed](8,16)--(12.9,7.46) node[midway,above right]{$d$};
      \fill(12.9,7.46) circle(.1) node[below right]{$C$};
    \end{tikzpicture}
  \end{center}
  \caption{Observing the steel balls while free-falling}
  \label{falling}
\end{figure}
 Using basic geometry, we can find distance $DE$ and $AE$:
 \begin{align*}
   DE&=8\tan\ang{30}=\SI{4.6}{\metre}\\
   AE&=16-DE=\SI{11.4}{\metre}
 \end{align*}
Using basic trigonometry again, we can now find the minimum distance of
separation:
\begin{displaymath}
  d=AE\cos\ang{30}=\boxed{\SI{9.9}{\metre}}
\end{displaymath}
The second part is slightly trickier, because we have (so far) ignored
acceleration due to gravity. However, there is no acceleration in the $x$
direction, i.e.\ the horizontal distance that $B$ travels is the same. We
need to compute $DC'$ which is just
\begin{displaymath}
  DC'=d\sin\ang{30}=\SI{4.9}{\metre}
\end{displaymath}
which means that in the time to reach minimum separation distance, the steel
ball $B$ has travelled $8+4.9=\SI{12.9}{\metre}$ horizontally. We can now use
the kinematic equation in the $x$-direction to compute $t$:
\begin{displaymath}
  t=\frac{\Delta x}{v_x}=\frac{12.9}{20\cos\ang{30}}=\boxed{\SI{.75}{\second}}
\end{displaymath}
Finally, we substitute $t$ back into the \emph{actual} position of $A$ and $B$
to compute their \emph{actual} location when the minimum separation occurs:
\begin{align*}
  \mb{x}_A &= 8\iii + (16-5t^2)\jjj = \boxed{8\iii+13.2\jjj} \\
  \mb{x}_B &=20\cos\ang{30}t\iii+(20\sin\ang{30}t-5t^2) \jjj
  =\boxed{12.9\iii+1.7\jjj}
\end{align*}
\end{document}
