\documentclass{../../oss-classkick-exam}

\begin{document}
\genheader

\gentitle{C}{GAUSS' LAW AND CAPACITORS}

\genmultidirections

\begin{questions}

    % THIS QUESTION IS TOO EASY; MOVE TO PHYSICS 2
%    \questions Two electric objects experience a repulsive force. What happens
%    to that force if the distance between the objects is doubled?
%    \begin{choices}
%      \choice It decreases to one-fourth its value.
%      \choice It decreases to one-half its value.
%      \choice It stays the same.
%      \choice It doubles.
%      \choice It quadruples.
%    \end{choices}
%    \vspace{.7in}

    % ALSO MOVE TO AP PHYSICS 2
%    \question A pith ball is a tiny piece of Styrofoam that is covered with a
%    conductive paint. One pith ball initially has a charge of \SI{6.4e-8}{C},
%    and it touches an identical, neutral pith ball. After the pith balls are
%    separated, what is the charge on the pith ball that had the initial charge?
%    \begin{choices}
%      \choice\SI{6.4e-8}{C}
%      \choice\SI{3.2e-8}{C}
%      \choice\SI{0}{C}
%      \choice\SI{-3.2e-8}{C}
%      \choice\SI{-6.4e-8}{C}
%    \end{choices}

    % ALSO MOVE TO AP PHYSICS 2
%    \question Glass becomes positively charged when it is rubbed with silk.
%    Which of the following is the best description of what’s happening?
%    \begin{choices}
%      \choice Electrons are rubbed off the glass onto the silk.
%      \choice Electrons are rubbed off the silk onto the glass.
%      \choice Protons are rubbed off the glass onto the silk.
%      \choice Protons are rubbed off the silk onto the glass.
%      \choice Neutrons in the glass have an affinity for positive charge.
%    \end{choices}

%    % ALSO MOVE TO AP PHYSICS 2
%    \question Consider an isolated, neutral system consisting of wool fabric
%    and a rubber rod. If the rubber rod is rubbed with wool to become negatively
%    charged, what can be said about the wool fabric?
%    \begin{choices}
%      \choice It becomes equally negatively charged.
%      \choice It becomes equally positively charged.
%      \choice It becomes negatively charged but not equally.
%      \choice It becomes positively charged but not equally.
%      \choice In a neutral system, neither object can become charged.
%    \end{choices}

  \question A non-conducting sphere does not have a uniform charge density,
  but the density $\rho$ varies with the distance $r$ from the center of the
  sphere according to the equation $\rho=\beta r$ where $\beta$ is a positive
  constant. The electric field inside the sphere ($r<R$) at a distance $r$
  from the center of the sphere is
  \begin{choices}
    \choice $\dfrac{\beta r^2}{12\epsilon_0}$
    \choice $\dfrac{\beta r^3}{3\epsilon_0}$
    \choice $\dfrac{\beta r}{2\epsilon_0}$
    \choice $\dfrac{\beta r^2}{2\epsilon_0}$
    \choice $\dfrac{\beta r^2}{4\epsilon_0}$
  \end{choices}
 
  \question The electric potential at the surface of the sphere from the last
  question is
  \begin{choices}
    \choice $\dfrac{\beta R^3}{12\epsilon_0}$
    \choice $\dfrac{\beta R}{2\epsilon_0}$
    \choice $\dfrac{\beta R^3}{3\epsilon_0}$
    \choice $\dfrac{\beta R^2}{2\epsilon_0}$
    \choice $\dfrac{\beta R^2}{4\epsilon_0}$
  \end{choices}
   
  \question According to Gauss's law, the net electric flux passing through a
  closed surface is
  \begin{choices}
    \choice positive if the flux is entering the surface
    \choice negative if the flux is exiting the surface
    \choice positive if the net charge inside the surface is zero
    \choice negative if the net charge inside the surface is zero
    \choice zero if the net charge inside the surface is zero
  \end{choices}
   
  \question According to Gauss's law, which of the following statements is
  true?
  \begin{choices}
    \choice It is possible to have a nonzero electric field, but zero electric
    flux.
    \choice It is possible to have a nonzero electric flux, but zero electric
    field.
    \choice It is possible to have a nonzero electric flux through a closed
    surface even if the enclosed charge in a surface is zero.
    \choice If a surface is not closed (such as a sheet of paper), the flux
    through it must be zero.
    \choice It is possible for charges located outside a closed surface to
    produce a net positive flux through the surface.
  \end{choices}
   
  \question Gauss's law is most convenient to use when calculating an electric
  field due to
  \begin{choices}
    \choice charges outside a closed surface
    \choice charges inside a closed surface that has high symmetry
    \choice charges inside a closed surface that has low symmetry
    \choice a potential difference that is negative
    \choice a potential difference that is positive
  \end{choices}

  \uplevel{
    \textbf{Question \ref{cube1}-\ref{cube2}}
    \cpic{.3}{cube}
  }
  \question A cube has sides of length $a$. The cube rests so that one side
  rests on the $x$-axis as shown. An electric field is established in the
  $x$-direction according to the function $E_x=bx^2$ , where $b$ is a positive
  constant. Which of the following statements is true?
  \label{cube1}
  \begin{choices}
    \choice There is a net charge inside the cube.
    \choice There is no net charge inside the cube.
    \choice The flux passing through the cube is negative.
    \choice The flux passing through the cube is zero.
    \choice The flux diminishes while passing through the cube.
  \end{choices}
   
  \question The charge inside the cube can be expressed by the equation
  \label{cube2}
  \begin{choices}
    \choice $\epsilon_0ba$
    \choice $\epsilon_0ba^2$
    \choice $\epsilon_0ba^3$
    \choice $\epsilon_0ba^4$
    \choice $\epsilon_0b^22a^2$
  \end{choices}

  \uplevel{
    \textbf{Questions \ref{cap1}--\ref{cap2}}
    \begin{center}
      \begin{tikzpicture}[scale=.5]
        \draw(0,0) circle(1);
        \draw(0,0) circle(3);
        \fill(0,0) circle(.1);
        \draw[->](0,0)--(1,0) node[midway,above]{a};
        \draw[->,rotate=-140](0,0)--(3,0) node[midway,left]{b};
        \begin{scope}[rotate=160]
          \draw[->](0,0)--(2,0) node[midway,above]{r};
          \fill(2,0) circle(.1);
        \end{scope}
        \node at (0,3.3) (a){$+Q$};
        \node at (0,1.3) (b){$-Q$};
      \end{tikzpicture}
    \end{center}
    The spherical capacitor shown above consists of a conducting shell of
    radius $a$ inside a larger conducting shell of radius $b$. A charge $-Q$
    is placed on the inner sphere and a charge $+Q$ is placed on the outer
    shell. The capacitance of the capacitor is $C_0$.
  }
  \question The magnitude of the electric field $E$ at a distance $r$ between
  the spheres is
  \label{cap1}
  \begin{choices}
    \choice $\dfrac Q{4\pi\epsilon_0r^2}$
    \choice $\dfrac Q{4\pi\epsilon_0r}$
    \choice $\dfrac Q{4\pi\epsilon_0a^2}$
    \choice $\dfrac Q{4\pi\epsilon_0b^2}$
    \choice zero
  \end{choices}

  \question The bottom half of the space between the spheres is filled with
  oil of dielectric constant $\kappa=3$, creating two capacitors connected to
  each other. Which of the following is true of the two capacitors?
  \begin{center}
    \begin{tikzpicture}[scale=.5]
      \draw[fill=gray!60](-1,0)--(-3,0) arc(180:360:3)--(1,0) arc(0:-180:1);
      \draw(0,0) circle(1);
      \draw(0,0) circle(3);
      \fill(0,0) circle(.1);
      \draw[->,rotate=230](0,0)--(1,0) node[midway,above left]{a};
      \draw[->](0,0)--(3,0) node[pos=.7,above]{b};
      \node at (0,3.3) (a){$+Q$};
      \node at (0,1.3) (b){$-Q$};
    \end{tikzpicture}
  \end{center}
  \begin{choices}
    \choice They are connected in series.
    \choice They are connected in parallel.
    \choice The total capacitance has not changed.
    \choice The total capacitance of the spheres has decreased.
    \choice The total capacitance is now zero.
  \end{choices}
    
  \question With the bottom half of the space between the spheres having been
  filled with oil of dielectric constant $\kappa=3$, the new capacitance of
  the spheres is
  \label{cap2}
  \begin{choices}
    \choice zero
    \choice $C_0$
    \choice $2C_0$
    \choice $3C_0$
    \choice $4C_0$
  \end{choices}   
\end{questions}
\newpage

\genfreetitle{C}{ELECTROSTATICS \& CAPACITORS}{6}

\genfreedirections

% TAKEN FROM 2007 AP PHYSICS C EXAM FREE-RESPONSE QUESTION E&M 2
\cpic{.3}{charges}
\begin{questions}
  \question In the figure above, a nonconducting solid sphere of radius a with
  charge $+Q$ uniformly distributed throughout its volume is concentric with a
  nonconducting spherical shell of inner radius $2a$ and outer radius $3a$ that
  has a charge $-Q$ uniformly distributed throughout its volume. Express all
  answers in terms of the given quantities and fundamental constants.

  \begin{parts}
    \part Using Gauss's law, derive expressions for the magnitude of the
    electric field as a function of radius $r$ in the following regions.
    \begin{subparts}
      \subpart Within the solid sphere ($r<a$)
      \subpart Between the solid sphere and the spherical shell ($a<r<2a$)
      \subpart Within the spherical shell ($2a<r<3a$)
      \subpart Outside the spherical shell ($r>3a$)
    \end{subparts}
    
    \part What is the electric potential at the outer surface of the spherical
    shell ($r=3a$)? Explain your reasoning.
    
    \part Derive an expression for the electric potential difference $V_X-V_Y$
    between points $X$ and $Y$ shown in the figure.
  \end{parts}
  \newpage

  % TAKEN FROM 2010 AP PHYSICS C EXAM FREE-RESPONSE QUESTION E&M 1
  \uplevel{
    \cpic{.28}{circle}
  }
  \question A charge $+Q$ is uniformly distributed over a quarter circle of
  radius $R$, as shown above. Points $A$, $B$, and $C$ are located as shown,
  with $A$ and $C$ located symmetrically relative to the $x$-axis. Express all
  algebraic answers in terms of the given quantities and fundamental constants.
  \begin{parts}
    \part Rank the magnitude of the electric potential at points $A$, $B$, and
    $C$ from greatest to least, with number 1 being greatest. If two points have
    the same potential, give them the same ranking. Justify your rankings.

    \vspace{.1in}
    \underline{\hspace{.3in} } $V_A$ \hspace{.3in}
    \underline{\hspace{.3in} } $V_B$ \hspace{.3in}
    \underline{\hspace{.3in} } $V_C$
    
    \uplevel{
      Point $P$ is at the origin, as shown below, and is the center of curvature
      of the charge distribution.
      \cpic{.28}{circle2}
    }

    \part Determine an expression for the electric potential at point $P$ due to
    the charge $Q$.

    \part A positive point charge $q$ with mass $m$ is placed at point $P$ and
    released from rest. Derive an expression for the speed of the point charge
    when it is very far from the origin.    

    \part On the dot representing point $P$ below, indicate the direction of the
    electric field at point P due to the charge $Q$.
    \begin{center}
      \begin{tikzpicture}[scale=1.2]
        \draw[dashed,->](-1,0)--(1,0) node[right]{$x$};
        \draw[dashed,->](0,-1)--(0,1) node[above]{$y$};
        \fill(0,0) circle(.1);
      \end{tikzpicture}
    \end{center}
    
    \part Derive an expression for the magnitude of the electric field at point
    $P$.
  \end{parts}
  \newpage
  
  % TAKEN FROM 2003 AP PHYSICS C EXAM FREE-RESPONSE QUESTION E&M 1. THE
  % TIKZ FIGURE BELOW IS A FAIRLY GOOD REPRODUCTION OF THE ORIGINAL AP EXAM
  % DIAGRAM.
  \uplevel{
    \centering
    \begin{tikzpicture}[scale=.85]
      \filldraw[dashed,thick,
        even odd rule,
        inner color=black!70,
        outer color=gray!20] (0,0) circle (2);
      \draw[very thick,->,rotate=45](0,0)--(2,0) node[midway,below]{$R$};
      \fill(6,0) circle(.05) node[below]{$P$};
    \end{tikzpicture}
  }
  \question A spherical cloud of charge of radius $R$ contains a total charge
  $+Q$ with a nonuniform volume charge density that varies according to the
  equation
  \begin{align*}
    \rho(r)&= \rho_0\left( 1 -\dfrac rR \right)\text{ for }r\leq R\text{ and}\\
    \rho(r)&= 0\text{ for }r>R
  \end{align*}
  where $r$ is the distance from the center of the cloud. Express all algebraic
  answers in terms of $Q$, $R$, and fundamental constants.
  \begin{parts}
    \part Determine the following as a function of $r$ for $r>R$.
    \begin{subparts}
      \subpart The magnitude $E$ of the electric field
      \subpart The electric potential $V$
    \end{subparts}
    \part A proton is placed at point $P$ shown above and released. Describe its
    motion for a long time after its release.
    
    \part An electron of charge magnitude $e$ is now placed at point $P$, which
    is a distance $r$ from the center of the sphere, and released. Determine
    the kinetic energy of the electron as a function of $r$ as it strikes the
    cloud.
    
    \part Derive an expression for $\rho_0$.
    
    \part Determine the magnitude $E$ of the electric field as a function of $r$
    for $r\leq R$.
  \end{parts}
  \newpage

  % TAKEN FROM THE 2004 AP PHYSICS C FREE-RESPONSE QUESTION E&M 1. NO CHANGES TO
  % THE QUESTION HAVE BEEN MADE
  \uplevel{
    \cpic{.65}{nonconcentric}
  }
  \question The figure above left shows a hollow, infinite, cylindrical,
  uncharged conducting shell of inner radius $r_1$ and outer radius $r_2$. An
  infinite line charge of linear charge density $+\lambda$ is parallel to its
  axis but off center. An enlarged cross section of the cylindrical shell is
  shown above right.
  \begin{parts}
    \part On the cross section above right,
    \begin{subparts}
      \subpart sketch the electric field lines, if any, in each of regions I,
      II, and III and
      \subpart use $+$ and $-$ signs to indicate any charge induced on the
      conductor.
      \end{subparts}
    \part In the spaces below, rank the electric potentials at points $a$, $b$,
    $c$, $d$, and $e$ from highest to lowest (1 = highest potential). If two
    points are at the same potential, give them the same number.

    \vspace{.1in}
    \underline{\hspace{.2in}} $V_a$\hspace{.3in}
    \underline{\hspace{.2in}} $V_b$\hspace{.3in}
    \underline{\hspace{.2in}} $V_c$\hspace{.3in}
    \underline{\hspace{.2in}} $V_d$\hspace{.3in}
    \underline{\hspace{.2in}} $V_e$
    \uplevel{
      \cpic{.65}{concentric}
    }

    \part The shell is replaced by another cylindrical shell that has the same
    dimensions but is nonconducting and carries a uniform volume charge density
    $+\rho$. The infinite line charge, still of charge density $+\lambda$, is
    located at the center of the shell as shown above. Using Gauss's law,
    calculate the magnitude of the electric field as a function of the distance
    $r$ from the center of the shell for each of the following regions. Express
    your answers in terms of the given quantities and fundamental constants.
    \begin{subparts}
      \subpart $r < r_1$
      \subpart $r_1\leq r\leq r_2$
      \subpart $r > r_2$
    \end{subparts}
  \end{parts}
  
%\item Two positive charges $+q$ are on the $y$ axis at $y=+a$ and $y=-a$.
%  \begin{enumerate}
%  \item Show that the electric field on the $x$ axis is along the $x$ axis with
%    $E_x=2kqx(x^2+a^2)^{-3/2}$.
%  \item Show that near the origin, when $x\ll a$, $E_x\approx 2kqx/a^3$.
%  \item Show that for $x\gg a$, $E_x\approx 2kq/x^2$.
%  \item Explain why you should expect the result in (c) even before calculating
%    it.
%  \end{enumerate}
%  A bead of mass $m$ with a negative charge $-q$ slides along a thread that
%  runs along the $x$ axis.
%  \begin{enumerate}[resume]
%  \item Show that for small displacements $x\ll a$, the bead experiences a
%    restoring force that is proportional to $x$ and therefore undergoes
%    simple harmonic motion.
%  \item Find the period of the motion.
%  \end{enumerate}
%  %\vspace{\stretch{4}}
%  \newpage
  
%\item Using Gauss's law, find
%  \begin{enumerate}
%  \item the electric field strength inside and outside of a uniformly charged
%    hollow sphere of radius $R$ and surface charge density $\sigma$ (charge
%    per unit area).
%  \item the electric field inside and outside an infinitely long cyclindrical
%    shell of charge of radius $R$ with charge discibution $\sigma$ (charge
%    per unit area).
%  \item the electric field strength inside and outside a infinitely long solid
%    cylinder of radius $R$ carrying a linear uniform charge density $\rho$
%    (charge per unit volume).
%  \end{enumerate}
%  Hint: In all cases, think about where to put the Gaussian surface. Take
%  advantage of symmetry.
%  %\vspace{\stretch{1}}
%  \newpage

%\item For the circuit shown below, find
%  \begin{enumerate}[noitemsep]
%  \item The total equivalent capacitance between the terminals
%  \item The charge stored on each capacitor
%  \item the total stored charge
%  \end{enumerate}
%  \begin{tikzpicture}[scale=1.2,american voltages]
%    \draw[thick] (0,2) to[C=0.30<\micro\farad>,*-] (3,2)
%    to[C=1.0<\micro\farad>] (3,0) to[short,-*](0,0);
%    \draw[thick] (3,2)--(5,2) to[C=0.25<\micro\farad>,] (5,0)--(3,0);
%  \end{tikzpicture}
%  \vspace{\stretch{1}}
  
%\item A parallel-plate capacitor has a capacitance $C_0$ and plate separation
%  of $d$. To dielectric slabs of constants $\kappa_1$ and $\kappa_2$, each of
%  thickness $d/2$ and having the same area as the plates, are inserted between
%  the plates as shown in the figure below. When the free charge on the plates
%  are $Q$,
%  \begin{enumerate}
%  \item find the electric field in each of the dielectric
%  \item find the potential difference between the plates
%  \item show that the new capacitance is given by:
%    $\displaystyle C=\frac{\kappa_1\kappa_2}{\kappa_1+\kappa_2}C_0$
%  \end{enumerate}
%  \pic{.2}{stacked}
%  %\vspace{\stretch{1}}

%\item Several point charges produce the equipotential lines shown.
%  \begin{enumerate}[noitemsep]
%  \item At which point on the diagram is the magnitude of the electric field
%    greatest? Explain.
%  \item Points C and D are approximately \SI{.02}{\metre} apart. Point F is
%    halfway between points C and D. What is the electric field at point F?
%  \item A \SI{5.}{\micro\coulomb} point charge is moved from point C to point
%    E, then to point D by an external force. Determine the work done by the
%    external force.
%  \end{enumerate}
%  \pic{.45}{equipotentials}
%  \vspace{\stretch{1}}

%\item A potential is given by
%  \begin{displaymath}
%    V(x,y,z)=\frac{kQ}{\sqrt{(x-a)^2+y^2+z^2}}
%  \end{displaymath}
%  \begin{enumerate}[noitemsep]
%  \item Find the components $E_x$, $E_y$ and $E_z$ of the electric field by
%    differentiating this potential function.
%  \item What simple charge distribution might be responsible for this potential?
%  \end{enumerate}
%  \vspace{\stretch{1}}
\end{questions}
\end{document}
