\documentclass{../../../oss-apphys}

\begin{document}
\genheader

\gentitle{2}{LIGHT AND GEOMETRIC OPTICS}

\genmultidirections

\gengravity

\raggedcolumns
\begin{multicols*}{2}
  \begin{enumerate}[leftmargin=18pt]
  \item In a laboratory experiment, you shine a green laser past a strand of
    hair. This produces a light and dark pattern on a screen. You notice that
    the lab group next to you has produced a similar pattern on a screen, but
    the light and dark areas are spread farther apart. Which of the following
    could cause the light and dark pattern to spread? (Select two answers.)
    \begin{enumerate}[nosep=0pt,leftmargin=18pt,label=(\Alph*)]
    \item The second group used thinner hair.
    \item The second group is using a red laser.
    \item The second group had the screen closer to the hair.
    \item The second group held the laser farther from the hair.
    \end{enumerate}
    
  \item An observer can hear sound from around a corner but cannot see light
    from around the same corner. Which of the following helps to explain this
    phenomenon?
    \begin{enumerate}[nosep=0pt,leftmargin=18pt,label=(\Alph*)]
    \item Sound is a longitudinal wave, and light is an electromagnetic wave.
    \item Sound is a mechanical wave, and light is a transverse wave.
    \item Light travels at a speed much faster than that of sound.
    \item Light has a much smaller wavelength than sound.
    \end{enumerate}
    
  \item A mirror produces an upright image one-half the height of the object
    when the object is 12 cm from the mirror's surface. What is the focal
    length of the mirror?
    \begin{enumerate}[nosep=0pt,leftmargin=18pt,label=(\Alph*)]
    \item\SI{-12}{\centi\metre}
    \item\SI{-4}{\centi\metre}
    \item\SI{4}{\centi\metre}
    \item\SI{6}{\centi\metre}
    \end{enumerate}

  \item A light ray with a wavelength of $\lambda_w$ and a frequency of
    $f_w$ in water ($n=1.33$) is incident on glass ($n=1.61$). In the glass,
    the wavelength and frequency of the light is $\lambda_g$ and $f_g$. How do
    the values of wavelength and frequency of the ray of light in water compare
    to those in glass?
    \cpic{.48}{meter-stick}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $\lambda_w>\lambda_g$, and $f_w=f_g$
    \item $\lambda_w>\lambda_g$, and $f_w>f_g$
    \item $\lambda_w<\lambda_g$, and $f_w=f_g$
    \item $\lambda_w<\lambda_g$, and $f_w<f_g$
    \end{enumerate}

  \item An optics bench is set up on a meter stick, as shown in the figure. The
    light source is a candle placed at $x_0$ . The lens is located at $x_1$. The
    screen is moved until a sharp image appears at location $x_2$. The data is
    recorded in a table, the lens is moved to a new location ($x_1$), and the
    screen is adjusted until the image is sharp again. Which of the
    following procedures will allow a student to determine the focal length
    of the lens?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item Plot $x_2$ as a function of $x_0$. The focal length will be the
      vertical axis intercept.
    \item Plot ($x_2-x_1$) as a function of ($x_0-x_1$). The focal length will
      be the vertical axis intercept.
    \item Plot $1/x_2$ as a function of $1/x_0$. The focal length will be the
      inverse of the vertical axis intercept.
    \item Plot $1/(x_2-x_1)$ as a function of $1/(x_0-x_1)$. The focal length
      will be the inverse of the vertical axis intercept.
    \end{enumerate}

  \item A laser beam passes through a prism and produces a bright dot of light
    a distance of $x$ from the prism, as shown in the figure. Which of the
    following correctly explains the change in distance as the angle
    ($\theta$) of the prism is decreased?
    \cpic{.3}{triangle}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The distance increases because the angle on incidence increases.
    \item The distance increases because the angle of incidence decreases.
    \item The distance decreases because the angle on incidence increases.
    \item The distance decreases because the angle of incidence decreases.
    \end{enumerate}

  \item In the human eye, the distance from the lens to the retina, on which the
    image is focused, is 20 mm. A book is held 30 cm from the eye, and
    the focal length of the eye is 16 mm. How far from the retina does the
    image form, and what lens should be used to place the image directly
    on the retina?
    
    \begin{tabular}{cll}
      & \underline{Distance of image from retina}
      & \underline{Corrective lens} \\
      (A) & 3.1 mm in front of the retina & Concave lens \\
      (B) & 3.1 mm in front of the retina & Convex lens \\
      (C) & 14 mm behind the retina & Concave lens \\
      (D) & 14 mm behind the retina & Convex lens
    \end{tabular}
  \end{enumerate}
\end{multicols*}
\newpage

\genfreetitle{2}{LIGHT AND GEOMETRIC OPTICS}{6}

\genfreedirections

% TAKEN FROM THE 2001 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #4
\begin{center}
  \pic{.4}{glass1}
\end{center}
\begin{enumerate}[leftmargin=15pt]
\item In an experiment a beam of red light of wavelength 675 nm in air passes
  from glass into air, as shown above. The incident and refracted angles are
  $\theta_1$ and $\theta_2$, respectively. In the experiment, angle $\theta_2$
  is measured for various angles of incidence $\theta_1$, and the sines of the
  angles are used to obtain the line shown in the following graph.
  \begin{center}
    \pic{.6}{graph1}
  \end{center}
  \begin{enumerate}
  \item Assuming an index of refraction of 1.00 for air, use the graph to
    determine a value for the index of refraction of the glass for the red
    light. Explain how you obtained this value.
  \item For this red light, determine the following.
    \begin{enumerate}
    \item The frequency in air
    \item The speed in glass
    \item The wavelength in glass
    \end{enumerate}
    
  \item The index of refraction of this glass is 1.66 for violet light, which
    has wavelength 425 nm in air.
    \begin{enumerate}[noitemsep]
    \item Given the same incident angle $\theta_1$ , show on the ray diagram on
      the previous page how the refracted ray for the violet light would vary
      from the refracted ray already drawn for the red light.
    \item Sketch the graph of $\sin\theta_2$ versus $\sin\theta_1$ for the
      violet light on the figure on the previous page that shows the same graph
      already drawn for the red light.
    \end{enumerate}

  \item Determine the critical angle of incidence $\theta_c$for the violet
    light in the glass in order for total internal reflection to occur.
  \end{enumerate}
  \vspace{1.5in}
  
  % THIS IS TAKEN FROM THE 2005 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #4
\item Your teacher gives you a slide with two closely spaced slits on it. She
  also gives you a laser with a wavelength $\lambda=\SI{632}{\nano\metre}$. The
  laboratory task that you are assigned asks you to determine the spacing
  between the slits. These slits are so close together that you cannot measure
  their spacing with a typical measuring device.
  \begin{enumerate}[noitemsep]
  \item From the list below, select the additional equipment you will need to
    do your experiment by checking the line next to each item.
    \begin{multicols}{3}
      \underline{\hspace{.3in}} Meterstick\\
      \underline{\hspace{.3in}} Ruler\\
      \underline{\hspace{.3in}} Tape measure\\
      \underline{\hspace{.3in}} Light-intensity meter\\
      \columnbreak
      
      \underline{\hspace{.3in}} Large screen\\
      \underline{\hspace{.3in}} Paper\\
      \underline{\hspace{.3in}} Slide holder\\
      \underline{\hspace{.3in}} Stopwatch
    \end{multicols}
  \item Draw a labeled diagram of the experimental setup that you would use. On
    the diagram, use symbols to identify carefully what measurements you will
    need to make.
    \begin{center}
      \begin{tikzpicture}[scale=.45]
        \draw[thick,->](-10,0)--(10,0) node[right]{Position};
        \draw[thick,->](0,-7)--(0,7) node[right]{Position};
      \end{tikzpicture}
    \end{center}
  \item On the axes below, sketch a graph of intensity versus position that
    would be produced by your setup, assuming that the slits are very narrow
    compared to their separation.
  \end{enumerate}
  \vspace{2in}
  
  % THIS IS TAKEN FROM THE 2014 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #7
  \begin{center}
    \pic{.35}{film1}\\
    \underline{Note:} Figure not drawn to scale.
  \end{center}
\item A thin layer of transparent oil is placed on top of a transparent plate.
  The oil film is then illuminated by white light shining onto the oil's
  surface, as shown in the figure above. To an observer standing right next to
  the light source and looking straight down on the oil film, the oil film
  appears green, corresponding to a wavelength of 520 nm in air. The oil has an
  index of refraction of 1.52.
  \begin{enumerate}
  \item Determine the frequency of the green light in the air.
  \item Determine the frequency of the green light in the oil film.
  \item Calculate the wavelength of the green light in the oil film.
  \item The oil film thickness is half of the wavelength you found in part (c).
    Is the index of refraction of the plate greater than, less than, or equal
    to that of the oil?

    \vspace{.1in}
    \underline{\hspace{.3in}} Greater than\hspace{.2in}
    \underline{\hspace{.3in}} Less than\hspace{.2in}
    \underline{\hspace{.3in}} Equal to
    
    \vspace{.1in}Justify your answer.

    \begin{center}
      \pic{.7}{film2}\\
      \underline{Note:} Figure not drawn to scale.
    \end{center}
  \item As the observer starts moving to the right away from the light source,
    as shown in the figures above, the film appears to change color. Describe
    the color change and give an explanation for this phenomenon.
  \end{enumerate}
  \newpage

  % THIS IS TAKEN FROM THE 2008 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #6
  \cpic{.8}{converging-mirror}  
\item The figure above shows a converging mirror, its focal point $F$, its
  center of curvature $C$, and an object represented by the solid arrow.
  \begin{enumerate}
  \item On the figure above, draw a ray diagram showing at least two incident
    rays and the image formed by them.
  \item Is the image real or virtual?

    \vspace{.2in}
    \underline{\hspace{.4in}} Real\hspace{.3in}
    \underline{\hspace{.4in}} Virtual

    \vspace{.1in}Justify your answer.
  \item The focal length of this mirror is \SI{6.}{\centi\metre}, and the
    object is located \SI{8.}{\centi\metre} away from the mirror. Calculate the
    position of the image formed by the mirror. (Do NOT simply measure your ray
    diagram.)
  \item Suppose that the converging mirror is replaced by a diverging mirror
    with the same radius of curvature that is the same distance from the
    object, as shown below.
    \cpic{.8}{diverging-mirror}
    For this mirror, how does the size of the image compare with that of the
    object?

    \vspace{.2in}
    \underline{\hspace{.4in}} Larger than the object\hspace{.3in}
    \underline{\hspace{.4in}} Smaller than the object\hspace{.3in}
    \underline{\hspace{.4in}} The same size as the object
    Justify your answer.
  \end{enumerate}
  \newpage

  % THIS IS TAKEN FROM THE 2010 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #5
  \begin{center}
    \pic{.4}{prism}\\
    \underline{Note:} Figure not drawn to scale.
  \end{center}
\item As shown above, a beam of red light of wavelength \SI{6.65e-7}\metre in
  air is incident on a glass prism at an angle $\theta_1$ with the normal. The
  glass has index of refraction $n=1.65$ for the red light. When
  $\theta_1=\ang{40}$, the beam emerges on the other side of the prism at an
  angle $\theta_4=\ang{84}$.
  \begin{enumerate}
  \item Calculate the angle of refraction $\theta_2$ at the left side of the
    prism.
  \item Using the same prism, describe a change to the setup that would result
    in total internal reflection of the beam at the right side of the prism.
    Justify your answer.
  \item The incident beam is now perpendicular to the surface. The glass is
    coated with a thin film that has an index of refraction $n_f = 1.38$ to
    reduce the partial reflection of the beam at this angle.
    \begin{enumerate}
    \item Calculate the wavelength of the red light in the film.
    \item Calculate the minimum thickness of the film for which the intensity
      of the reflected red ray is near zero.
    \end{enumerate}
  \end{enumerate}
  \newpage

  % THIS IS TAKEN FROM THE 2006 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #4
\item A student performs an experiment to determine the index of refraction $n$
  of a rectangular glass slab in air. She is asked to use a laser beam to
  measure angles of incidence $\theta_i$ in air and corresponding angles of
  refraction $\theta_r$ in glass. The measurements of the angles for five
  trials are given in the table below.
  \begin{center}
    \bgroup
    \def\arraystretch{1.7}
    \begin{tabular}{|c|c|c|p{1in}|p{1in}|}
      \hline
      Trial & $\theta_i$ & $\theta_r$ & & \\ \hline
      \hspace{.2in}1\hspace{.2in} & \hspace{.2in}\ang{30}\hspace{.2in} &
      \hspace{.2in}\ang{20}\hspace{.2in} & & \\ \hline
      2 & \ang{40} & \ang{27} & & \\ \hline
      3 & \ang{50} & \ang{32} & & \\ \hline
      4 & \ang{60} & \ang{37} & & \\ \hline
      5 & \ang{70} & \ang{40} & & \\ \hline
    \end{tabular}
    \egroup
  \end{center}
  \begin{enumerate}
  \item Complete the last two columns in the table by calculating the
    quantities that need to be graphed to provide a linear relationship from
    which the index of refraction can be determined. Label the top of each
    column.
  \item On the grid below, plot the quantities calculated in (a) and draw an
    appropriate graph from which the index of refraction can be determined.
    Label the axes.
    
    \begin{center}
      \begin{tikzpicture}[scale=.85]
        \draw[thick](0,0) rectangle(14,10);
        \draw(0,0) grid(14,10);
        \foreach\x in {0,.2,.4,...,14}\draw[thick](\x,-.1)--(\x,.1);
        \foreach\y in {0,.2,.4,...,10}\draw[thick](-.1,\y)--(.1,\y);
      \end{tikzpicture}
    \end{center}
  \item Using the graph, calculate the index of refraction of the glass slab.
  \end{enumerate}
  \newpage
  \begin{center}
    \begin{tikzpicture}[scale=1.2]
      \fill(5,2) circle(.15) node[right]{\hspace{.1in}Light Source};
      \draw[very thick](5,2)--(5,0);
      \draw[very thick,->](5,2)--(5,.8);
      \draw[very thick](0,0)--(10,0);
      \fill[gray!50](0,-.2) rectangle(10,-.5);
      \draw[very thick](0,-.2)--(10,-.2);
      \node at (10,.2) [right]{Air};
      \node at (10,-.1)[right]{Oil Film};
      \node at (10,-.4)[right]{Water};
    \end{tikzpicture}
  \end{center}
  The student is also asked to determine the thickness of a film of oil
  ($n=1.43$) on the surface of water ($n=1.33$). Light from a variable
  wavelength source is incident vertically onto the oil film as shown above.
  The student measures a maximum in the intensity of the reflected light when
  the incident light has a wavelength of \SI{600}{\nano\metre}.
  \begin{enumerate}[resume]
  \item At which of the two interfaces does the light undergo a \ang{180} phase
    change on reflection?

    \vspace{.1in}
    \underline{\hspace{.4in}} The air-oil interface only\hspace{.3in}
    \underline{\hspace{.4in}} The oil-water interface only\hspace{.3in}
    \underline{\hspace{.4in}} Both interfaces\hspace{.3in}\\
    \underline{\hspace{.4in}} Neither interface
    \vspace{.1in}
    
  \item Calculate the minimum possible thickness of the oil film.
  \end{enumerate}
\end{enumerate}
\end{document}
