\documentclass[12pt,compress,aspectratio=169]{beamer}
\input{../mybeamer}
\usepackage{pgfplots}

\usetikzlibrary{decorations.pathmorphing,patterns}

\title{Classes 9 \& 10: Harmonic Motion}
\subtitle{Advanced Placement Physics C}
\input{../me}
\input{../mycommands}


\begin{document}

\begin{frame}
  \maketitle
\end{frame}


\section{Hooke's Law}

\begin{frame}{Review: Hooke's Law}
  Hooke's law for an ideal spring relates the \textbf{spring force} $\vec F_s$
  exerted by a compressed or stretched spring onto another object to the
  \textbf{spring constant} $k$\footnote{It is the stiffness of the spring, also
  called \textbf{Hooke's constant}, \textbf{force constant}, or the
  \textbf{spring rate} of the spring. It depends on both the geometry of the
  spring, as well as the material properties of the spring}, and the spring's
  displacement $\vec x$:

  \eq{-.1in}{
    \boxed{\vec F_s=-k\vec x}
  }
  
  \begin{center}
    \begin{tabular}{l|c|c}
      \rowcolor{pink}
      \textbf{Quantity} & \textbf{Symbol} & \textbf{SI Unit} \\ \hline
      Spring force        & $\vec F_s$ & \si\newton \\
      Spring constant     & $k$        & \si{\newton\per\metre}\\
      Spring Displacement & $\vec x$   & \si\metre
    \end{tabular}
  \end{center}
  \vspace{.3in}
\end{frame}



\begin{frame}{Review: Elastic Potential Energy}
  Applying Hooke's law in the work equation gives the amount of \textbf{elastic
    potential energy} stored in the spring when it is compressed or stretched:

  \eq{-.1in}{
    W=\int^{x_1}_{x_0}F_s\dl x=-\int^{x_1}_{x_0}kx\dl x
    =-\frac12kx^2\Big|^{x_1}_{x_0}=-\Delta U_s
  }

  where elastic potential energy is defined as

  \eq{-.1in}{
    \boxed{U_s=\frac12kx^2}
  }
\end{frame}



\section{Spring-Mass Systems}

\begin{frame}{Horizontal Spring-Mass System}
  Consider the forces acting on a mass connected horizontally to a spring
  without friction
  \begin{center}
    \begin{tikzpicture}[scale=1.1]
      \draw[mass] (5,.5) rectangle (6,1.5);
      \draw[thick,decorate,
        decoration={aspect=.4,segment length=8,amplitude=8,coil}] (0,1)--(5,1);
      \fill[pattern=north east lines] (6.5,.5)--(6.5,.3)--(-0.2,.3)
      --(-.2,2)--(0,2)--(0,.5)--cycle;
      \draw[thick] (6.5,.5)--(0,.5)--(0,2);
      \draw[axes] (7,1)--(7.8,1) node[right]{$+$};
      \draw[dashed] (3.5,-.5)--(3.5,2.2) node[above]{Equilibrium};
      \draw[vectors] (3.5,0)--(5,0) node[midway,below]{$x$};
      \fill[red] (5.5,1) circle (.08);
      \draw[vectors,red] (5.5,1)--(5.5,0) node[below]{$\vec F_g$};
      \draw[vectors,red] (5.5,1)--(5.5,2) node[above]{$\vec F_N$};
      \draw[vectors,red] (5.5,1)--(4.5,1) node[above]{$\vec F_s$};
    \end{tikzpicture}
  \end{center}
  $\vec F_g$ and $\vec F_N$ cancel out, so net force is due only to spring force
  $\vec F_s=-k\vec x$ along the $x$-axis. This is true both when the spring is
  in compression or extension.\footnote{It should be obvious that the spring is
  in extension in the diagram.}
  \vspace{.2in}
\end{frame}



\begin{frame}{Horizontal Spring-Mass System}
  Applying second law of motion in the $x$-direction:

  \eq{-.1in}{
    \sum F=F_s=ma\quad\longrightarrow\quad-kx=m\diff[2]xt
  }

  This is a
  \emph{second-order ordinary differential equation with constant
    coefficients}. In standard form:

  \eq{-.1in}{
    \diff[2]xt + \frac kmx=0
  }
\end{frame}



\begin{frame}{Horizontal Spring-Mass System}

  The solution to the equation is a function $x(t)$ where the second time
  derivative $\ddot x$ looks like $x$ but with a negative sign

  \eq{-.1in}{
    \diff[2]xt + \frac kmx=0
  }

  The obvious choices are the trigonometric functions: $\sin(t)$ and $\cos(t)$.
  Starting with this general form:
    
  \eq{-.1in}{
    x(t)=A\cos(\omega_0 t-\phi)
  }
  
  Cosine is \emph{usually} preferred because $\cos(0)=1$, and since
  oscillations often begin at amplitude at $t=0$, therefore $\phi=0$.
  Mathematically, the two functions only differ in $\phi$
\end{frame}



\begin{frame}{Horizontal Spring-Mass System}

  \eq{0in}{
    \boxed{\diff[2]xt + \frac kmx=0}
  }
  
  Starting with the general form and take the time derivatives to obtain the
  velocity and acceleration of the mass:
 
  \vspace{-.2in}{\large
    \begin{align*}
      x(t)&=A\cos(\omega_0 t-\phi)\\
      v(t)&=-A\omega_0\sin(\omega_0 t-\phi)\\
      a(t)&=-A\omega_0^2\cos(\omega_0 t-\phi)=-\omega_0^2x
    \end{align*}
  }
  
  $\omega_0$ is the angular frequency of the oscillation, $A$ is the amplitude,
  and $\phi$ is the phase shift, based on initial condition
\end{frame}



\begin{frame}{Horizontal Spring-Mass System: Angular Frequency}
  Substituting expressions of $x(t)$ and $a(t)=\ddot x$ back into the ODE, we
  find that the solution is satisfied if $\omega_0$ is related to the spring
  constant and mass by:

  \eq{-.1in}{
    \boxed{
      \omega_0=\sqrt{\frac km}
    }  }

  The angular frequency for the simple harmonic oscillator is called the
  \textbf{natural frequency}. The period ($T_0$) and frequency ($f_0$) of the
  simple harmonic oscillator are then given by:

  \eq{-.1in}{
    f_0=\frac{\omega_0}{2\pi}=\frac1{2\pi}\sqrt{\frac km}\quad\quad
    T_0=\frac1{f_0}=2\pi\sqrt{\frac mk}
  }
  
  Angular frequency, frequency, and period do not depend on amplitude $A$
\end{frame}



\begin{frame}{Displacement, Velocity and Acceleration}
  \begin{columns}
    \column{.4\textwidth}
    \begin{align*}
      x(t)&=A\cos(\omega_0 t-\phi)\\
      v(t)&=-A\omega_0\sin(\omega_0 t-\phi)\\
      a(t)&=-A\omega_0^2\cos(\omega_0 t-\phi)=-\omega_0^2x
    \end{align*}

    \column{.6\textwidth}
    \begin{tikzpicture}[yscale=.5]
      \draw[axes] (0,-2)--(0,2) node[left]{$x$};
      \draw[axes] (-1,0)--(8,0) node[right]{$t$};
      \draw[functions,smooth,samples=80,domain=0:7.5]
      plot(\x,{1.5*cos(90*\x)});
      
      \draw[axes] (0,-7)--(0,-3)  node[left]{$v$};
      \draw[axes] (-1,-5)--(8,-5) node[right]{$t$};
      \draw[blue,very thick,smooth,samples=80,domain=0:7.5]
      plot(\x,{-1.5*sin(90*\x)-5});
      
      \draw[axes] (0,-12)--(0,-8)  node[left]{$a$};
      \draw[axes] (-1,-10)--(8,-10) node[right]{$t$};
      \draw[violet,very thick,smooth,samples=80,domain=0:7.5]
      plot(\x,{-1.5*cos(90*\x)-10});
    \end{tikzpicture}
  \end{columns}
\end{frame}



\begin{frame}{Side Note \#1}
  \textbf{Side Note \#1:} For anyone with more experiences with calculus,
  you will know that the solution to any ODE is the linear combination of
  \emph{all} possible solutions\footnote{From an even broader discussions on
  mathematics, all the possible solutions to the ODE must be functions that are
  \emph{orthogonal} to each other.}, i.e.:

  \eq{-.1in}{
    x(t)=c_1\sin(\omega t)+c_2\cos(\omega t)
  }

  \vspace{-.1in}Where $c_1$ and $c_2$ are coefficients based on initial
  position $x_0$ and velocity $v_0$.\footnote{Trigonometric identities can be
  used to show that the solution form shown in previous slides is identical.}
\end{frame}



\begin{frame}{Side Note \#2}
  \textbf{Side Note \#2:} Another function that we can be try is the
  exponential function, where its derivatives is related to the function
  itself. However, the second derivative of $e^{\omega t}$ does not have the
  negative sign that is needed to solve the problem.

  \eq{-.1in}{
    x(t)=e^{\omega t}\quad\rightarrow\quad
    \diff[2]xt=\omega^2 e^{\omega t}=\omega^2x
  }

  But if the exponential function is \emph{imaginary}:

  \eq{-.1in}{
    x(t)=e^{i\omega t}
  }

  \vspace{-.1in}Then\ldots
\end{frame}



\begin{frame}{Side Note \#2}
  The second derivative \emph{does} in fact have the negative sign:

  \vspace{-.3in}{\large
    \begin{align*}
      x&=e^{i\omega t}\\
      \dot x&=i\omega e^{i\omega t}\\
      \ddot x&=i^2\omega^2 e^{i\omega t}=-\omega^2 e^{i\omega t}
    \end{align*}
  }
  
  This should not come as a surprise, since the complex exponential function
  and the sinusoidal functions are related:

  \eq{-.1in}{
    e^{i\omega t}=\cos(\omega t)+i\sin(\omega t)
  }
\end{frame}



\begin{frame}{Vertical Spring-Mass System}
  \begin{columns}
    \column{.3\textwidth}
    \begin{tikzpicture}[scale=1.3]
      \draw[mass] (.7,1.7) rectangle (1.3,2.3);
      \draw[thick,
        decoration={aspect=0.3,segment length=2mm, amplitude=2.5mm, coil},
        decorate] (1,5)--(1,2.5); 
      \fill[pattern=north east lines] (0,5) rectangle (2,5.2);
      \draw[very thick] (0,5)--(2,5);
      \draw[vectors,red] (1,2)--(1,1.2) node[right]{$\vec F_g$};
      \draw[vectors,red] (1,2)--(1,2.8) node[right]{$\vec F_s$};
      \fill[red] (1,2) circle (.05);
      \draw[axes] (1,1)--(1,.5) node[below]{$+$};
      \draw[vectors] (.3,4)--(.3,2.3) node[midway,left]{$x$};
      \draw[vectors] (1.7,4)--(1.7,3.5) node[midway,right]{$B$};
      \begin{scope}[thick,dashed]
        \draw (0,4)--(2,4) node[right]{\scriptsize unstretched};
        \draw (0,3.5)--(2,3.5) node[right]{\scriptsize equilibrium};
      \end{scope}
    \end{tikzpicture}

    \column{.7\textwidth}
    For a vertical spring-mass system, we must consider
    {\color{purple}weight $F_g=mg$} as well:

    \eq{-.1in}{
      {\color{purple}F_g}-kx=m\diff[2]xt
    }

    But since $F_g$ is constant, the only change is the addition of a
    constant $B$ in the expression of $x(t)$:
    
    \vspace{-.25in}{\large
      \begin{align*}
        x(t) &= A\cos(\omega_0 t-\phi) +B\\
        v(t) &= -A\omega_0\sin(\omega_0 t-\phi)\\
        a(t) &= -A\omega_0^2\cos(\omega_0 t-\phi)
      \end{align*}
    }
  \end{columns}
\end{frame}



\begin{frame}{Vertical Spring-Mass System}
  \begin{columns}
    \column{.3\textwidth}
    \begin{tikzpicture}[scale=1.3]
      \draw[mass] (.7,1.7) rectangle (1.3,2.3);
      \draw[thick,
        decoration={aspect=0.3,segment length=2mm, amplitude=2.5mm, coil},
        decorate] (1,5)--(1,2.5); 
      \fill[pattern=north east lines] (0,5) rectangle (2,5.2);
      \draw[very thick] (0,5)--(2,5);
      \draw[vectors,red] (1,2)--(1,1.2) node[right]{$\vec F_g$};
      \draw[vectors,red] (1,2)--(1,2.8) node[right]{$\vec F_s$};
      \fill[red] (1,2) circle (.05);
      \draw[axes] (1,1)--(1,.5) node[below]{$+$};
      \draw[vectors] (.3,4)--(.3,2.3) node[midway,left]{$x$};
      \draw[vectors] (1.7,4)--(1.7,3.5) node[midway,right]{$B$};
      \begin{scope}[thick,dashed]
        \draw (0,4)--(2,4) node[right]{\scriptsize unstretched};
        \draw (0,3.5)--(2,3.5) node[right]{\scriptsize equilibrium};
      \end{scope}
    \end{tikzpicture}

    \column{.7\textwidth}
    $B$ is found by substituting $x$ and $\ddot x$ into the ODE. It is the
    stretch of the spring due to its own weight:
    
    \eq{-.1in}{
      B=\frac{mg}k
    }

    Angular frequency (natural frequency) remains the same as the horizontal
    case:

    \eq{-.1in}{
      \omega_0=\sqrt{\frac km}
    }
  \end{columns}
\end{frame}



\begin{frame}{Conservation of Energy in a Spring-Mass System}

  In the spring-mass systems, if there are no frictional losses, then the only
  forces doing work are the spring force (horizontal and vertical) and gravity
  (vertical). Both forces are \emph{conservative}, therefore the total
  mechanical energy is conserved:

  \eq{-.1in}{
    K + U_s + U_g = K' + U_s' + U_g'
  }
  
  For the horizontal spring-mass system, the total energy of the simple harmonic
  oscillator is:
    
  \eq{-.1in}{
    \boxed{E_T=\frac12kA^2}
  }
\end{frame}



\begin{frame}{Simple Example}
  \textbf{Example:} A mass suspended from a spring is oscillating up and
  down. Consider the following two statements:
  \begin{enumerate}
  \item At some point during the oscillation, the mass has zero velocity but it
    is accelerating
  \item At some point during the oscillation, the mass has zero velocity and
    non-zero acceleration.
  \end{enumerate}

  \begin{enumerate}[(A)]
  \item Both occur at some time during the oscillation
  \item Neither occurs during the oscillation
  \item Only (1) occurs
  \item Only (2) occurs
  \end{enumerate}
\end{frame}



\begin{frame}{Another Example}
  \textbf{Example:} An object of mass \SI5{\kilo\gram} hangs from a spring
  and oscillates with a period of \SI{.5}\second. By how much will the
  equilibrium length of the spring be shortened when the object is removed.
  \begin{enumerate}[(A)]
  \item\SI{.75}{\centi\metre}
  \item\SI{1.5}{\centi\metre}
  \item\SI{3.1}{\centi\metre}
  \item\SI{6.2}{\centi\metre}
  \end{enumerate}
\end{frame}



\section{Simple Pendulum}

\begin{frame}{What About a Simple Pendulum?}
  \begin{columns}[T]
    \column{.75\textwidth}
    \begin{itemize}
    \item Pendulums also exhibit oscillatory motion
    \item In a \textbf{simple pendulum}, all of the mass is concentrated at the
      end point 
    \item Two forces act on the mass: weight $\vec F_g$ and tension $\vec F_T$
    \item It has been shown in Class 6 (circular motion) that when the mass is
      deflected by an angle $\theta$, the tangential force is
      $F_\text{tangent}=-mg\sin\theta$
    \item No need to worry about the radial direction; it does not have to do
      with the restoring force
    \end{itemize}

    \column{.25\textwidth}
    \centering 
    \begin{tikzpicture}
      \fill[pattern=north east lines] (-1,0) rectangle (1,0.2);
      \draw[thick] (-1,0)--(1,0);
      \begin{scope}[rotate=20]
        \draw[thick] (0,0)--(0,-5) node[midway,right]{$\ell$};
        \shade[ball color=red] (0,-5) circle (.2) node[below right]{$m$};
        \draw[vectors,red,dotted] (0,-5)--(-1.5*sin{20},-5)
        node[left,fill=yellow!10]{$mg\sin\theta$};
        \draw[vectors,red] (0,-5)--(0,-3) node[left]{$\vec F_T$};
        \draw[vectors,red,rotate around={-20:(0,-5)}] (0,-5)--(0,-6.5)
        node[below]{$\vec F_g$};
      \end{scope}
      \draw[axes] (0,-2) arc (270:290:2) node[midway,below]{$\theta$};
      \draw[dashed] (0,0)--(0,-5);
      \draw[dashed] (0,-5) arc (270:295:5);
      \draw[dashed] (0,-5) arc (270:255:5);
    \end{tikzpicture}
  \end{columns}
\end{frame}



\begin{frame}{The Simple Pendulum}
  \begin{columns}
    \column{.77\textwidth}
    Substitute $F_\text{tangent}$ into the second law of motion, and cancelling
    $m$:

    \eq{-.2in}{
      F_\text{tangent}=ma_\text{tangent}\quad\rightarrow\quad
      -mg\sin\theta=m\ell\diff[2]{\theta}t
    }

    \vspace{-.15in}Solving this ODE is difficult because of the $\sin\theta$
    term. However, we can simplify the problem by using the Taylor series
    expansion of the sine function:

    \eq{-.1in}{
      \sin\theta
      =\theta-\frac{\theta^3}{3!}+\frac{\theta^5}{5!}-\frac{\theta^7}{7!}+\cdots
    }
    
    which shows that for small angles, $\sin\theta\approx\theta$
    
    \column{.22\textwidth}
    \centering
    \begin{tikzpicture}
      \fill[pattern=north east lines] (-1,0) rectangle (1,0.2);
      \draw[thick] (-1,0)--(1,0);
      \begin{scope}[rotate=20]
        \draw[thick] (0,0)--(0,-5) node[midway,right]{$\ell$};
        \shade[ball color=red] (0,-5) circle (.2) node[below right]{$m$};
      \end{scope}
      \draw[axes] (0,-2) arc (270:290:2) node[midway,below]{$\theta$};
      \draw[dashed] (0,0)--(0,-5);
      \draw[dashed] (0,-5) arc (270:295:5);
      \draw[dashed] (0,-5) arc (270:255:5);
    \end{tikzpicture}
  \end{columns}
\end{frame}



\begin{frame}{The Simple Pendulum}
  \begin{columns}
    \column{.77\textwidth}
    For small angles of $\theta$, the ODE reduces to the same form as the
    spring-mass system

    \eq{-.1in}{
      \diff[2]{\theta}t + \frac g\ell\theta=0
    }

    So how small is ``small angle''? That depends on what tolerance (the
    number of significant figures) is needed in the answer.
    
    \column{.22\textwidth}
    \begin{tikzpicture}
      \fill[pattern=north east lines] (-1,0) rectangle (1,0.2);
      \draw[thick] (-1,0)--(1,0);
      \begin{scope}[rotate=20]
        \draw[thick] (0,0)--(0,-5) node[midway,right]{$\ell$};
        \shade[ball color=red] (0,-5) circle (.2) node[below right]{$m$};
      \end{scope}
      \draw[axes] (0,-2) arc (270:290:2) node[midway,below]{$\theta$};
      \draw[dashed] (0,0)--(0,-5);
      \draw[dashed] (0,-5) arc (270:295:5);
      \draw[dashed] (0,-5) arc (270:255:5);
    \end{tikzpicture}
  \end{columns}
\end{frame}



\begin{frame}{Ordinary Differential Equation for the Pendulum}
  The solution for $\theta(t)$ is a sinusoidal function, like the spring-mass
  system:

  \eq{-.1in}{
    \boxed{\theta(t)=\theta_\text{max}\cos(\omega_0 t-\phi)}
  }

  where $\theta_\text{max}$ is the maximum deflection (amplitude), and $\phi$
  is a phase shift based on the initial condition of the pendulum. (If the
  oscillation begins at amplitude, then $\phi=0$) The natural frequency of
  the oscillation $\omega_0$ is given by:
    
  \eq{-.1in}{
    \boxed{
      \omega_0=\sqrt{\frac g\ell}
    }
  }
\end{frame}



\begin{frame}{How Good is the Small Angle Approximation?}
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
          width=2.5in,
          xmin=0,xmax=pi/2,
          ymin=0,ymax=pi/2,
          %xlabel=$\theta$ (radian),
          xtick={0,pi/8,pi/4,3*pi/8,pi/2},
          xticklabels={
            0,$\dfrac\pi8$,$\dfrac\pi4$,$\dfrac{3\pi}8$,$\dfrac\pi2$
          },
          grid = both,
          legend pos=north west,
        ]
        \addplot[
          color=blue,
          domain=0:pi/2,
          samples=40,
          style={thick}]{sin(x*180/pi)};
        \addlegendentry{$\sin\theta$}
        \addplot[
          color=red,
          domain=0:pi/2,
          samples=40,
          thick]{x};
        \addlegendentry{$\theta$}
      \end{axis}
    \end{tikzpicture}
    \begin{tikzpicture}
      \begin{axis}[
          width=2.5in,
          ylabel=\% Error,
          xmin=0,xmax=pi/2,
          ymin=0,ymax=60,
          %xlabel=$\theta$ (radian),
          xtick={0,pi/8,pi/4,3*pi/8,pi/2},
          xticklabels={
            0,$\dfrac\pi8$,$\dfrac\pi4$,$\dfrac{3\pi}8$,$\dfrac\pi2$
          },
          ytick={0,10,...,60},
          legend pos=north west,
          grid = both,
        ]
        \addplot[
          color=violet,
          domain=0:pi/2,
          samples=40,
          style=thick
        ]{abs(sin(x*180/pi)-x)/sin(x*180/pi)*100};
      \end{axis}
    \end{tikzpicture}
  \end{center}
  What the maximum angle of deflection can be used depends on the precision
  that the answer requires.
\end{frame}



\begin{frame}{Pendulum Example Problem}
  \textbf{Example:} A simple pendulum consists of a mass $m$ attached to a
  light string of length $\ell$. If the system is oscillating through small
  angles, which of the following is true
  \begin{enumerate}[(A)]
  \item The frequency is independent of the acceleration due to gravity, $g$.
  \item The period depends on the amplitude of the oscillation.
  \item The period is independent of the mass $m$.
  \item The period is independent of the length $\ell$.
  \end{enumerate}
\end{frame}



\begin{frame}{A Pendulum Example}
  \textbf{Example:} A bucket full of water is attached to a rope and allowed
  to swing back and forth as a pendulum from a fixed support. The bucket has a
  hole in its bottom that allows water to leak out. How does the period of
  motion change of the bucket with the loss of water?
  \begin{enumerate}[(A)]
  \item The period does not change.
  \item The period continuously decreases.
  \item The period continuously increases.
  \item The period increases to some maximum and then decreases again.
  \end{enumerate}
\end{frame}



\begin{frame}{Think About $g$}
  \textbf{Example:} A little girl is playing with a toy pendulum while riding
  in an elevator. Being an astute and educated young lass, she notes that the 
  period of the pendulum is $T=\SI{.5}\second$. Suddenly the cables
  supporting the elevator break and all  of the brakes and safety features fail
  simultaneously. The elevator plunges into free fall. The young girl is
  astonished to discover that the pendulum has:
  \begin{enumerate}[(A)]
  \item continued oscillating with a period of \SI{.5}\second.
  \item stopped oscillating entirely.
  \item decreased its rate of oscillation to have a longer period.
  \item increased its rate of oscillation to have a lesser period.
  \end{enumerate}
\end{frame}



\section{Damped Oscillation}

\begin{frame}{It's Never Perfect}
  In reality, there are friction, or drag, or other damping forces present in
  the spring-mass system, represented schematically by the shock absorber:
  \begin{center}
    \begin{tikzpicture}
      \draw[gray,mass] (3,.5) rectangle (4,1.5);
      \draw[thick,draw=gray,decorate,
        decoration={aspect=.4,segment length=5,amplitude=8,coil}] (0,1)--(3,1);
      \fill[gray,pattern=north east lines] (-.2,2)--(-.2,.3)--(6.7,.3)--(6.7,2)
      --(6.5,2)--(6.5,.5)--(0,.5)--(0,2)--cycle;
      \draw[gray,thick] (0,2)--(0,.5)--(6.5,.5)--(6.5,2);
      \draw[axes] (7.25,1)--(8,1) node[right]{$x$};
      \fill[blue!20!white] (5,.8) rectangle (6,1.2);
      \draw[very thick] (4,1)--(5,1);
      \draw[very thick] (5,.8)--(5,1.2);
      \draw[very thick] (4.9,1.2)--(6,1.2)--(6,.8)--(4.9,.8);
      \draw[very thick] (6,1)--(6.5,1);
      \uncover<2->{
        \draw[vectors,pink] (3.5,1)--(3.5,0) node[right]{$\vec F_g$};
        \draw[vectors,pink] (3.5,1)--(3.5,2) node[right]{$\vec F_N$};
        \draw[vectors,pink] (3.5,1)--(2.5,1) node[above]{$\vec F_s$};
        \draw[vectors,red] (3.5,1)--(4.5,1) node[above]{$\vec F_D$};
        \fill[red] (3.5,1) circle (.075);
      }
    \end{tikzpicture}
  \end{center}
  \uncover<3->{
    \vspace{-.1in}The damping force is typically related to velocity, in the
    opposite direction:

    \eq{-.135in}{
      \vec F_D=-b(\vec v)^n
    }

    \vspace{-.12in}where $b$ is a positive constant called the \textbf{damping
      factor}. The simplest case is to use $n=1$ to represent combined viscous
    effects. %(Kinetic friction, $n=0$; aerodynamic drag, $n=2$)
  }
\end{frame}



\begin{frame}{Damped Oscillator}
  The 2nd-order ODE is obtained by applying second law of motion, this time
  with the additional term from the damping force:

  \eq{-.1in}{
    \sum F=F_s+{\color{red}F_D}=ma\quad\rightarrow\quad
    -kx{\color{red}-b\diff xt}=m\diff[2]xt
  }

  Arranging into standard form:
  
  \eq{-.1in}{
    \diff[2]xt+\frac bm\diff xt+\frac kmx=0
  }

  The solution to this ODE is still relatively straightforward (but not as
  easy).
\end{frame}



\begin{frame}{Damped Oscillator}
  The solution to this ODE\footnote{This is still a standard problem in
    calculus.} has both an {\color{magenta}exponential decay term} and a
  {\color{cyan}sinusoidal term}:

  \eq{-.1in}{
    x(t)=\underbrace{A_0{\color{magenta}e^{-\frac b{2m}t}}}_\text{amplitude}
    {\color{cyan}\cos(\omega t+\phi)}
  }

  where $A_0$ is the initial amplitude of the damped oscillator,
  and the ``natural frequency'' for the damped oscillator is given by:

  \eq{-.1in}{
    \omega=\sqrt{\omega_0^2-\left(\frac b{2m}\right)^2}
  }
  
  Note that $\omega<\omega_0$ (natural frequency decreases) because of the
  damping factor $b$.
  \vspace{.3in}
\end{frame}



\begin{frame}{Critical Damping}
  \textbf{Critical damping} occurs when the natural frequency $\omega$ is zero,
  i.e.:
  
  \eq{-.1in}{
    \sqrt{\omega_0^2-\left(\frac b{2m}\right)^2}=0
  }

  which corresponds to the \textbf{critical damping constant} $b_c$:

  \eq{-.1in}{
    b_c=2m\omega_0%=2\sqrt{km}
  }
  \begin{itemize}
  \item A critically damped system returns to its equilibrium position in the
    shortest time with \emph{no} oscillation
  \item When $b>b_c$, the system is \textbf{over-damped}
  \item Critical or near-critical damping is desired in many engineering
    designs (e.g.\ shock absorbers on car suspensions)
  \end{itemize}
\end{frame}



\begin{frame}{Comparing Damped System}
  \centering
  \pic{.5}{oscda8}
  
  The motion of the damped oscillator is not strictly periodic.
\end{frame}



%\begin{frame}{Energy in a Damped System}
%  The non-conservative damping force dissipates energy from the oscillator at
%  a rate of:
%
%  \eq{-.1in}{
%    P=\diff Et=\vec F_D\cdot\vec v=-bv^2
%  }
%
%  As velocity relate to energy by: $(v_{av})^2=E/m$, power dissipation is a
%  first-order linear ODE:
%
%  \eq{-.1in}{
%    \diff Et=-\frac bmE
%  }
%
%  The solution to the ODE shows the total amount of energy decreases
%  exponentially with time:
%
%  \eq{-.1in}{
%    E(t)=E_0e^{-\frac bmt} %=E_0e^{-\frac{t}{\tau}}
%    }
%\end{frame}



\section{Driven Oscillation}

\begin{frame}{Forced Harmonic Motion}
  To keep a damped system going, energy must be added into the system. Assuming
  that the system is subjected to an external applied force ($F_a$) that is
  harmonic with time, with a driving frequency $\omega_a$:

  \eq{-.2in}{
    F_a=F\cos(\omega_at)
  }
  
  \vspace{-.2in}
  \begin{center}
    \begin{tikzpicture}
      \draw[gray,mass] (3,.5) rectangle (4,1.5);
      \draw[thick,gray,decorate,
        decoration={aspect=.4,segment length=6,amplitude=8,coil}] (0,1)--(3,1);
      \fill[gray,pattern=north east lines] (-.5,2)--(-.5,.3)--(6.7,.3)--(6.7,2)
      --(6.5,2)--(6.5,.5)--(0,.5)--(0,2)--cycle;
      \draw[gray,thick] (0,2)--(0,.5)--(6.5,.5)--(6.5,2);
      \draw[fill=blue!10] (4.9,.8) rectangle (5.5,1.2);
      \draw[very thick,gray] (4,1)--(4.9,1);
      \draw[very thick,gray] (4.9,.8)--(4.9,1.2);
      \draw[very thick,gray] (4.8,1.2)--(5.5,1.2)--(5.5,.8)--(4.8,.8);
      \draw[very thick,gray] (5.5,1)--(6.5,1);
      \draw[vectors,pink] (3.5,1)--(3.5,0) node[right]{$\vec F_g$};
      \draw[vectors,pink] (3.5,1)--(3.5,2) node[right]{$\vec F_N$};
      \draw[vectors,pink] (3.5,1)--(2.5,1) node[above]{$\vec F_s$};
      \draw[vectors,red] (3.5,.95)--(2,.95) node[below]{$\vec F_a$};
      \draw[vectors,pink] (3.5,1)--(4.5,1) node[below]{$\vec F_D$};
      \fill[red] (3.5,1) circle (.075);
    \end{tikzpicture}
  \end{center}
  In general, the driving frequency $\omega_a$ is unrelated to the undamped
  natural frequency $\omega_0$ or damped frequency $\omega$.
\end{frame}



\begin{frame}{Forced Harmonic Motion}
  Again, the second-order ordinary differential equation is obtained by
  applying the second law of motion:
  
  \eq{-.1in}{
    \sum F=-kx-bv+F\cos(\omega_at)=ma
  }

  Rearranging the terms gives a similar ODE to the damped case, but with the
  {\color{magenta}additional applied external force term} on the right-hand
  side:

  \eq{-.1in}{
    m\diff[2]xt+b\diff xt+kx={\color{magenta}F\cos(\omega_at)}
  }
\end{frame}



\begin{frame}{Forced Harmonic Motion}
  This is a much more difficult problem in calculus, but nevertheless still a
  standard problem
  
  \eq{-.1in}{
    m\diff[2]xt + b\diff xt+kx=F\cos(\omega_a t)
  }
  
  The solution to this ODE has two components:
  \begin{itemize}
  \item A \textbf{transient solution} that is identical to that of the damped
    oscillator
    \begin{itemize}
    \item Obtained by setting $F_a=0$
    \item Depends on the initial condition
    \item Solution becomes negligible over time because of exponential-decay
    \end{itemize}
  \item A \textbf{steady-state solution} which does not depend on the initial
    condition
  \end{itemize}
\end{frame}



\begin{frame}{Forced Harmonic Motion}
  Solving for the steady-state solution will be left as a difficult calculus
  exercise\footnote{This is usually taught in a 2nd-year university level ODE
  course}, but it can be shown that the solution is a harmonic motion at the
  driving frequency $\omega_a$ of the external force:

  \eq{-.1in}{
    \boxed{x(t)=A\cos(\omega_a t-\phi)}
  }
  
  where the amplitude of the oscillation $A$ and phase shift $\phi$ are given
  by:

  \eq{-.05in}{
    A=\frac F{\sqrt{m^2(\omega_0^2-\omega_a^2)^2+b^2\omega_a^2}}
    \quad\;\;
    \tan\phi=\frac{b\omega_a}{m(\omega_0^2-\omega_a^2)}
  }
  \vspace{.5in}
\end{frame}



\begin{frame}{Amplitude Response to External Driving Force}
  The amplitude of the driven oscillation is given by:
  
  \eq{-.1in}{
    A=\frac F{\sqrt{m^2(\omega_0^2-\omega_a^2)^2+b^2\omega_a^2}}
  }
  
  Maximum $A$ occurs when the denominator is minimized, i.e.\ when the
  derivative with respect to external frequency $\omega_a$ is 0:

  \eq{-.25in}{
    \diff{}{\omega_a}\left[m^2(\omega_0^2-\omega_a^2)^2+b^2\omega_a^2\right]=0
    \quad\longrightarrow\quad
    \omega_a=\sqrt{\omega_0^2-\frac{b^2}{2m^2}}\approx\omega
  }

  The driving frequency $\omega_a$ when amplitude is maximum occurs at a
  frequency that is very close to (albeit slightly different from) the natural
  frequency of the damped system.
\end{frame}



\begin{frame}{Resonance}
  \begin{columns}
    \column{.5\textwidth}
    \begin{tikzpicture}[scale=.95]
      \begin{axis}[
          width=3.3in,
          xmin=0,xmax=3, xlabel=Driving Frequency ($\omega_a$),
          xtick={0,1,2,3}, xticklabels={0,$\omega$,$2\omega$,$3\omega$},
          ymin=0,ymax=6, ylabel=Relative Amplitude ($A$),
          ytick={0,1,2,3,4,5,6}
        ]
        \addplot[
          color=red!25,
          samples=50,
          domain=0:3,
          very thick]{1/sqrt((1-x^2)^2+x^2)};
        \addlegendentry{$\frac b{2m}=\frac12\omega$}
        \addplot[
          color=red!50,
          samples=80,
          domain=0:3,
          very thick]{1/sqrt((1-x^2)^2+x^2/4)};
        \addlegendentry{$\frac b{2m}=\frac14\omega$}
        \addplot[
          color=red!75,
          samples=250,
          domain=0:3,
          very thick]{1/sqrt((1-x^2)^2+x^2/16)};
        \addlegendentry{$\frac b{2m}=\frac18\omega$}
        \addplot[
          color=red,
          samples=400,
          domain=0:3,
          very thick]{1/sqrt((1-x^2)^2+x^2/32)};
        \addlegendentry{$\frac b{2m}=\frac1{16}\omega$}
      \end{axis}
    \end{tikzpicture}

    \column{.5\textwidth}
    Plotting amplitude $A$ as a function of driving frequency $\omega_a$ shows
    that:
    \begin{itemize}
    \item For a lightly damped system (i.e.\ small $b$), resonance response is
      highest when $\omega_a\approx\omega\approx\omega_0$
    \item The smaller the damping constant $b$, the higher and narrower the
      peak is
    \end{itemize}
  \end{columns}
\end{frame}



\begin{frame}{Resonance}
  \eq{0in}{
    \boxed{\tan\phi=\frac{b\omega_a}{m(\omega_0^2-\omega_a^2)}}
  }

  When $\omega_E=\omega_0$ is substituted into the phase shift expression, the
  right-hand side becomes undefined. From this, we obtain a phase shift of
  $\phi=\pi/2$. Taking derivative of $x(t)$ for velocity $v(t)$, and
  substituting $\phi=\pi/2$:
  
  \eq{-.1in}{
    v(t)=\dot x
    =-A\omega_a\sin(\omega_a t-\frac{\pi}2)
    =A\omega_a\cos(\omega_a t)
  }
\end{frame}



\begin{frame}{Resonance}
  At resonance, the object is always moving in the same direction as the
  driving force:

  \vspace{-.3in}{\large
    \begin{align*}
      v(t)&=A\omega_a\cos(\omega_a t)\\
      F_a(t)&=F\cos(\omega_a t)
    \end{align*}
  }

  This also makes sense from a work-energy perspective, because now the
  external force is doing positive work to the system.
\end{frame}



\begin{frame}{Resonance}
  \textbf{Resonance} is caused by in-phase excitation near the natural
  frequency. This means that:
  \begin{itemize}
  \item The frequency of the driving force is \emph{approximately} equal to the
    natural frequency of the damped oscillator:

    \eq{-.1in}{
      \omega_a=\sqrt{\omega_0^2-\frac{b^2}{2m^2}}
      \quad\text{where}\quad\omega_0=\sqrt{\frac km}
    }

    For a lightly damped system, $\omega_a\approx\omega\approx\omega_0$
  \item The driving force follows the motion of the oscillator.
  \end{itemize}
\end{frame}
\end{document}
