\documentclass{../../../oss-classkick}

\begin{document}
\genheader

\gentitle{1}{CIRCULAR MOTION AND GRAVITY}

\genmultidirections

\gengravity

\raggedcolumns
\begin{multicols}{2}

  \begin{enumerate}[leftmargin=18pt]

  \item A \SI{1000}{\kilo\gram} car experiences a centripetal force of
    \SI{1.8e5}{\newton} while making a turn. The car is moving at a constant
    speed of \SI{30}{\metre\per\second}. What is the radius of the turn?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{.2}{\metre}
    \item\SI{1}{\metre}
    \item\SI{2}{\metre}
    \item\SI{4}{\metre}
    \item\SI{5}{\metre}
    \end{enumerate}

  \item A record player has four coins at different distances from the center of
    rotation. Coin A is 1 cm away, Coin B is 2 cm away, Coin C is 4 cm away,
    and Coin D is 8 cm away. If the player is spinning 45 rotations/min, what
    coin has the greatest tangential velocity?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item Coin A
    \item Coin B
    \item Coin C
    \item Coin D
    \item All the coins have equal tangential velocities.
    \end{enumerate}
    \vspace{.7in}
    
  \item Friction allows a car to make a turn at a speed of 10 miles per hour. By
    what factor will the friction have to change to allow the driver to make the
    same turn at twice the speed?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item Four times the friction
    \item Twice the friction
    \item The same amount of friction
    \item One-half the friction
    \item One-fourth the friction
    \end{enumerate}
    \vspace{.7in}
    
  \item A tetherball is whirled in a horizontal circle above your head. If the
    string breaks, the ball will follow what type of path if it is observed
    from above?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item Straight outward from the center
    \item Straight toward the center
    \item An expanding spiral
    \item A curved path that gradually approaches a straight line
    \item Tangent to the original circular path
    \end{enumerate}
    \vspace{.7in}
    
  \item A pendulum bob is attached to a string that is tied to the ceiling, and
    the bob is pulled back and released. As the bob moves through the bottom of
    the swing, how does the magnitude of the tension force from the string
    compare to the gravitational force on the bob?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The tension force is less than the gravitational force.
    \item The tension force is greater than the gravitational force.
    \item The tension force is equal to the gravitational force.
    \item The mass of the ball is needed in order to compare these forces.
    \item The release height of the ball is needed in order to compare these
      forces.
    \end{enumerate}
    \vspace{.7in}
    \columnbreak
    
  \item A pendulum bob is attached to a string that is tied to the ceiling, and
    the bob is pulled back and released from different heights. As the bob moves
    through the bottom of the swing, how is its centripetal acceleration related
    to its speed?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The centripetal acceleration is directly proportional to the speed
      of the pendulum.
    \item The centripetal acceleration is inversely proportional to the speed of
      the pendulum.
    \item  The centripetal acceleration is directly proportional to the square
      of the speed of the pendulum.
    \item The centripetal acceleration is inversely proportional to the square
      of the speed of the pendulum.
    \item There is no relationship between the centripetal acceleration and
      the speed.
    \end{enumerate}
    \vspace{.7in}
    
  \item Two satellites of equal mass orbit a planet. Satellite B orbits at twice
    the orbital radius of Satellite A. Which of the following statements is
    true?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The gravitational force on Satellite A is four times less than that on
      Satellite B.
    \item The gravitational force on Satellite A is two times less than that on
      Satellite B.
    \item The gravitational force on the satellites is equal.
    \item The gravitational force on Satellite A is two times greater than that
      on Satellite B.
    \item The gravitational force on Satellite A is four times greater than that
      on Satellite B.
    \end{enumerate}
    \vspace{.7in}
    
  \item A proposed ``space elevator'' can lift a \SI{1000}{\kilo\gram} payload
    to an orbit of \SI{150}{\kilo\metre} above the Earth's surface. The radius
    of the Earth is \SI{6.4e6}{\metre}, and the Earth's mass is
    \SI{6.e24}{\kilo\gram}. What is the gravitational potential energy of the
    payload when it reaches orbit?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{1.0e3}{\joule}
    \item\SI{2.7e6}{\joule}
    \item\SI{6.1e10}{\joule}
    \item\SI{2.7e12}{\joule}
    \item\SI{1.0e15}{\joule}
    \end{enumerate}
  \end{enumerate}
  \columnbreak
  
  \textbf{Questions \ref{centrifugal1}--\ref{centrifugal2}} are based on the
  following graph:
  \cpic{.45}{a-graph}

  \begin{enumerate}[leftmargin=18pt,resume]
  \item Engineers have designed a centrifuge for studying the effects of high
    gravity environments on plants and animals. This graph shows the
    results of the relationship between the radius and the centripetal
    acceleration. If the scientists want to simulate a “3-G environment,”
    then what should the radius of the centrifuge be?
    \label{centrifugal1}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item 1 m
    \item 2 m
    \item 3 m
    \item 5 m
    \item 10 m
    \end{enumerate}
    
  \item If an astronaut with a mass of 70 kg were placed in that centrifuge with
    a radius of 5 m, what would be the centripetal force acting on him?
    \label{centrifugal2}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item 30 N
    \item 70 N
    \item 140 N
    \item 210 N
    \item 240 N
    \end{enumerate}
    
  \item A satellite orbits the Earth at a distance of \SI{200}{\kilo\metre}. If
    the mass of the Earth is \SI{6.e24}{\kilo\gram} and the Earth's radius is
    \SI{6.4e6}{\metre}, what is the satellite's speed?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{1.e3}{\metre\per\second}
    \item\SI{3.5e3}{\metre\per\second}
    \item\SI{7.8e3}{\metre\per\second}
    \item\SI{5e6}{\metre\per\second}
    \item\SI{6.1e7}{\metre\per\second}
    \end{enumerate}
    
  \item Mars orbits the Sun at a distance of \SI{2.3e11}{\metre}. The mass of
    the Sun is \SI{2.e30}{\kilo\gram}, and the mass of Mars is
    \SI{6.4e23}{\kilo\gram}. Approximately what is the gravitational force that
    the Sun exerts on Mars?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{1.6e20}{\newton}
    \item\SI{1.6e21}{\newton}
    \item\SI{3.7e21}{\newton}
    \item\SI{3.7e32}{\newton}
    \item\SI{3.7e42}{\newton}
    \end{enumerate}
    
  \item When climbing from sea level to the top of Mount Everest, a hiker
    changes elevation by \SI{8848}{\metre}. By what percentage will the
    gravitational field of the Earth change during the climb? (The Earth's
    mass is \SI{6.e24}{\kilo\gram}, and its radius is \SI{6.4e6}{\metre}.)
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item It will increase by approximately \SI{.3}{\percent}.
    \item It will decrease by approximately \SI{.3}{\percent}.
    \item It will increase by approximately \SI{12}{\percent}.
    \item It will decrease by approximately \SI{12}{\percent}.
    \item The gravitational field strength will not change.
    \end{enumerate}
    \vspace{.7in}
    
  \item Four planets, A through D, orbit the same star. The relative masses and
    distances from the star for each planet are shown in the table. For
    example, Planet A has twice the mass of Planet B, and Planet D has
    three times the orbital radius of Planet A. Which planet has the highest
    gravitational attraction to the star?
    \begin{center}
      \vspace{-.1in}
      \begin{tabular}{lll}
        \hline
        \textbf{Planet} & \textbf{Relative mass} & \textbf{Relative distance}\\
        \hline
        A\hspace{.4in}& $2m$     & $r$    \\ \hline
        B & $m$                  & $0.1r$\hspace{.25in} \\ \hline
        C & $0.5m$\hspace{.25in} & $2r$   \\ \hline
        D & $4m$                 & $3r$   \\ \hline
      \end{tabular}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item Planet A
    \item Planet B
    \item Planet C
    \item Planet D
    \item All have the same gravitational attraction to the star.
    \end{enumerate}
    \columnbreak
    
  \item A satellite orbits the Earth at a distance that is four times the radius
    of the Earth. If the acceleration due to gravity near the surface of the
    Earth is $g$, the acceleration of the satellite is most nearly
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item zero
    \item $g/2$
    \item $g/4$
    \item $g/8$
    \item $g/16$
    \end{enumerate}

  \item The mass of a planet is 1/4 that of Earth and its radius is half of
    Earth's radius. The acceleration due to gravity on this planet is most
    nearly
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{2}{\metre\per\second\squared}
    \item\SI{4}{\metre\per\second\squared}
    \item\SI{5}{\metre\per\second\squared}
    \item\SI{10}{\metre\per\second\squared}
    \item\SI{20}{\metre\per\second\squared}
    \end{enumerate}
  
  \item Two planets of mass $M$ and $9M$ are in the same solar system. The
    radius of the planet of mass $M$ is $R$. In order for the acceleration due
    to gravity to be the same for each planet, the radius of the planet of mass
    $9M$ would have to be
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $R/2$
    \item $R$
    \item $2R$
    \item $3R$
    \item $9R$
    \end{enumerate}
    
  \item Two planets, X and Y, orbit a star. Planet X orbits at a radius $R$, and
    Planet Y orbits at a radius $3R$. Which of the following best represents
    the relationship between the acceleration $a_X$ of Planet X and the
    acceleration $a_Y$ of Planet Y?
    \begin{center}
      \begin{tikzpicture}[scale=.6]
        \tikzstyle{balloon}=[ball color=gray];
        \shade[balloon](0,0) circle(.25);
        \draw[dashed](0,0) circle(1);
        \draw[dashed](0,0) circle(3);
        \draw(0,0)--(1,0) node[pos=1,right]{X} node[midway,above]{$R$};
        \draw[fill=black](1,0) circle(.05);
        \begin{scope}[rotate=230]
          \draw(0,0)--(3,0) node[pos=1,left]{Y} node[pos=0.7,right]{$3R$};
          \draw[fill=black](3,0) circle(.05);
        \end{scope}
      \end{tikzpicture}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $a_X = 9a_Y$
    \item $9a_X = a_Y$
    \item $a_X = 3a_Y$
    \item $3a_X = a_Y$
    \item $a_X = a_Y$
    \end{enumerate}
    
  \item A satellite is in a stable circular orbit around the Earth at a radius
    $R$ and speed $v$. At what radius would the satellite travel in a stable
    orbit with a speed $2v$?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $R/4$
    \item $R/2$
    \item $R$
    \item $2R$
    \item $4R$
    \end{enumerate}

  \item The Earth and the moon apply a gravitational force to each other. Which
    of the following statements is true?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The Earth applies a greater force on the moon than the moon exerts on
      the Earth.
    \item The Earth applies a smaller force on the moon than the moon exerts on
      the Earth.
    \item The Earth applies a force on the moon, but the moon does not exert a
      force on the Earth.
    \item The Earth does not apply a force on the moon, but the moon exerts a
      force on the Earth.
    \item The force the Earth applies to the moon is equal and opposite to the
      force the moon applies to the Earth.
    \end{enumerate}
    \vspace{.7in}
    \columnbreak
    
  \item Two masses exert a gravitational force $F$ on each other. If one of the
    masses is doubled, and the distance between the masses is tripled, the
    new force between them is
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $6F$
    \item $2F/3$
    \item $2F/9$
    \item $3F/2$
    \item $4F/9$
    \end{enumerate}
    
  \item A planet orbits at a radius $R$ around a star of mass $M$. The period of
    orbit of the planet is
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $\displaystyle\sqrt{\frac{4\pi^2R^2}{GM}}$
    \item $\displaystyle\frac{4\pi^2R^3}{GM}$
    \item $\displaystyle\sqrt{\frac{4\pi^2R^3}{GM}}$
    \item $\displaystyle\sqrt{\frac{4\pi^2R}{GM}}$
    \item $\displaystyle\frac{GM}{4\pi^2R}$
    \end{enumerate}
    \columnbreak
    
  \item If a planet has twice the radius of Earth and half of Earth's density,
    what is the acceleration due to gravity on the surface of the planet (in
    terms of the gravitational acceleration $g$ on the surface of Earth)?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $4g$
    \item $2g$
    \item $g$
    \item $g/2$
    \item $g/4$
    \end{enumerate}

  \item The graph below depicts the tangential velocities of several circular
    space stations with different radii. All the stations are spinning. Which
    of the following statements is true?
    \cpic{.44}{stations}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The centripetal accelerations of the three shorter radii space
      stations are greater than \SI{10}{\metre\per\second\squared}; those of
      the larger ones are less than \SI{10}{\metre\per\second\squared}.
    \item  The centripetal accelerations of the three shorter radii space
      stations are greater than \SI{5}{\metre\per\second\squared}; those of
      the larger ones are less than \SI{5}{\metre\per\second\squared}.
    \item The centripetal accelerations of all the stations are all nearly
      \SI{5}{\metre\per\second\squared}.
    \item The centripetal accelerations of all the stations are all nearly
      \SI{10}{\metre\per\second\squared}.
    \item The centripetal accelerations of the three shorter radii space
      stations are less than \SI{10}{\metre\per\second\squared}; those of the
      larger ones are greater than \SI{10}{\metre\per\second\squared}.
    \end{enumerate}
  \end{enumerate}
\end{multicols}
\newpage

\genfreetitle{1}{CIRCULAR MOTION AND GRAVITY}{5}

\genfreedirections

% THIS QUESTION IS FROM THE 2001 AP PHYSICS B FREE RESPONSE QUESTION #1
\begin{center}
  \begin{tikzpicture}[scale=.5]
    \draw[dashed](0,0) circle(3);
    \fill(-3,0) circle(.15) node[left]{$P$};
    \fill(0,-3) circle(.15) node[below]{$Q$};
    \fill(0,3) circle(.15) node[above]{$Z$};
    \draw[very thick,rotate=50,->](4.5,0) arc(0:-30:4.5);
    \begin{scope}[rotate=35]
      \draw[thick](0,0)--(3,0) node[pos=.6,left]{$R$};
      \fill(3,0) circle(.3) node[above right]{$M$};
    \end{scope}
  \end{tikzpicture}
\end{center}
\begin{enumerate}[leftmargin=15pt]
\item A ball of mass $M$ is attached to a string of length $R$ and negligible
  mass. The ball moves clockwise in a vertical circle, as shown above. When the
  ball is at point $P$, the string is horizontal. Point $Q$ is at the bottom of
  the circle and point $Z$ is at the top of the circle. Air resistance is
  negligible. Express all algebraic answers in terms of the given quantities
  and fundamental constants.
  \begin{enumerate}[nosep]
  \item On the figures below, draw and label all the forces exerted on the ball
    when it is at points $P$ and $Q$, respectively.
    \begin{center}
      \begin{tikzpicture}[scale=.6]
        \draw[dashed](0,0) circle(3);
        \fill(-3,0) circle(.25) node[left]{$P\;$};
        \draw[very thick,rotate=-75,->](3.5,0) arc(0:-30:3.5);
      \end{tikzpicture}
      \hspace{1in}
      \begin{tikzpicture}[scale=.6]
        \draw[dashed](0,0) circle(3);
        \fill(0,-3) circle(.25) node[above]{$Q$};
        \draw[very thick,rotate=-30,->](3.5,0) arc(0:-30:3.5);
      \end{tikzpicture}
    \end{center}
    \vspace{.3in}
  \item Derive an expression for $v_\mathrm{min}$, the minimum speed the ball
    can have at point $Z$ without leaving the circular path.
  \item The maximum tension the string can have without breaking is
    $T_\mathrm{max}$. Derive an expression for $v_\mathrm{min}$, the maximum
    speed the ball can have at point $Q$ without breaking the string.
  \item Suppose that the string breaks at the instant the ball is at point $P$.
    Describe the motion of the ball immediately after the string breaks.
  \end{enumerate}
  \newpage
  
\item Two stars of equal mass $M$ are orbiting each other in a circular path.
  Show that the orbital period is given by:
  \begin{displaymath}
    T^2=\frac{2\pi^2d^3}{GM}
  \end{displaymath}
  where $d$ is the distance between the stars.
  \vspace{\stretch{1}}
  
\item Two stars of unequal mass orbit each other about their common center of
  mass as shown. The star of mass $M_1$ orbits in a circle of radius $r$, and
  the star of mass $M_2$ orbits in a circle of radius $2r$.
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \tikzstyle{balloon}=[ball color=gray];
      \draw[fill=black](0,0) circle(.075);
      \draw[dashed](0,0) circle(3);
      \draw[dashed](0,0) circle(1.5);
      \draw[->](0,0)--(1.5,0) node[midway,below]{$r$};
      \shade[balloon] (1.5,0) circle (0.2) node[right]{$M_1$};
      \draw[->](0,0)--(-3,0) node[midway,below]{$2r$};
      \shade[balloon] (-3,0) circle (0.2) node[left]{$M_2$};
    \end{tikzpicture}
  \end{center}
  \begin{enumerate}[nosep]
  \item Determine the ratio of masses $M_1/M_2$.
  \item Determine the ratio of the acceleration $a_1$ of $M_1$ to the
    acceleration $a_2$ of $M_2$.
  \item Determine the ratio of the period $T_1$ of $M _1$ to the period $T_2$
    of $M_2$.
  \end{enumerate}
  \vspace{\stretch{1}}
  \newpage

\item Two point particles of mass $m$ are on the $y$ axis at $y=a$ and $y=-a$,
  as shown in the figure below.
  \begin{center}
    \begin{tikzpicture}[scale=.8]
      \tikzstyle{balloon1}=[ball color=green!50!gray];
      \tikzstyle{balloon2}=[ball color=red!50];
      \draw[->](-2,0)--(6,0) node[pos=1,right]{$x$};
      \draw[->](0,-3)--(0,3) node[pos=1,above]{$y$};
      \draw[<->](-1,0)--(-1,2) node[midway,left]{$a$};
      \draw[<->](-1,0)--(-1,-2)node[midway,left]{$a$};
      \draw(0,2)--(-1.2,2);
      \draw(0,-2)--(-1.2,-2);
      \shade[balloon1] (0,2) circle (.55) node[white]{$m$};
      \shade[balloon1] (0,-2)circle (.55) node[white]{$m$};
      \shade[balloon2] (4,0) circle (.45) node[white]{$m_0$};
    \end{tikzpicture}
  \end{center}
  \begin{enumerate}[nosep]
  \item Derive the expression for the gravitational force exerted by these two
    particles on a third particle of mass $m_0$ located on the $x$ axis at a
    distance $x$ away from the origin.
  \item What is the gravitational field $\mb{g}$ on the $x$-axis due to the
    two particles?
  \item Show that $g_x$ (the $x$ component of $\mb{g}$) due to the two
    particles on the $y$ axis is approximately $\displaystyle-\frac{2Gm}{x^2}$
    when $x$ is much greater than $a$.
  \item Show that the maximum value of $|g_x|$ occurs at the point
    $\displaystyle x=\frac{\pm a}{\sqrt{2}}$.
  \end{enumerate}
  \newpage

\item Five equal masses $M$ are equally spaced on the arc of a semicircle of
  radius $R$ as shown in the figure below. A mass $m$ is located at the center
  of curvature of the arc. If $M$ is \SI{3}{\kilo\gram}, $m$ is
  \SI{21}{\kilo\gram}, and $R$ is \SI{10}{\centi\metre}, what is the force on
  $m$ due to the five masses?
  \begin{center}
    \begin{tikzpicture}[scale=.8]
      \tikzstyle{balloon1}=[ball color=gray!50];
      \tikzstyle{balloon2}=[ball color=red!60!black];
      \draw[->](0,0)--(5,0) node[pos=1,right]{$x$};
      \draw[->](0,0)--(0,5) node[pos=1,above]{$y$};
      \draw[very thick,blue!80](4,0) arc(0:180:4);
      \shade[balloon1] (0,0) circle (.1) node[below]{$m$};
      \foreach \theta in {0,45,...,180} {
        \begin{scope}[rotate=\theta]
          \shade[balloon2] (4,0) circle (.3) node[white]{$M$};
        \end{scope}
      }
    \end{tikzpicture}
  \end{center}
  \vspace{\stretch{1}}
\end{enumerate}
\end{document}
