\documentclass{../../oss-apphys-exam}
\newcounter{lastmc}

\begin{document}
\genheader

\gentitle{7}{ROTATIONAL MOTION, PART 1}

%\genmultidirections
%\gengravity

\raggedcolumns
\begin{multicols*}{2}
  \begin{questions}

    \question A meter stick of mass \SI{.1}{\kilo\gram} rests on a table as
    shown. A length of \SI{40}{\centi\metre} extends over the edge of the table.
    How far from the edge of the table could a \SI{.05}{\kilo\gram} mass be
    placed on the meter stick so that the stick just begins to tip?
    \begin{center}
      \begin{tikzpicture}[scale=4.5]
        \fill[lightgray](0,-.25) rectangle(.6,0);
        \begin{scope}[thick]
          \draw rectangle(1,.06);
          \draw(.6,0)--(.6,-.25);
          \draw(.75,.06) rectangle (.85,.14);
        \end{scope}
        \draw[|<->|](.6,-.05)--(1,-.05) node[midway,below]{40 cm};
        \fill(.5,.03) circle(.015);
        \node at (.5,.13){C};
      \end{tikzpicture}
    \end{center}
    \begin{choices}
      \choice\SI5{\centi\metre}
      \choice\SI{10}{\centi\metre}
      \choice\SI{15}{\centi\metre}
      \choice\SI{20}{\centi\metre}
      \choice\SI{30}{\centi\metre}
    \end{choices}
    
    \question A metal bar of constant density and weight $W$ is attached to a
    pivot on the wall at point $P$ and supported by a rope that makes an angle
    of \ang{60} with the vertical wall. The reaction force exerted by the pivot
    on the bar at point $P$ is best represented by which arrow?
    \cpic{.25}{metal-bar}
    \begin{choices}
      \choice{\Huge$\nearrow$}
      \choice{\Huge$\uparrow$}
      \choice{\Huge$\downarrow$}
      \choice{\Huge$\nwarrow$}
      \choice{\Huge$\searrow$}
    \end{choices}
    \columnbreak
    
    \question A ballet dancer is spinning around a vertical axis with her arms
    fully extended. How are her angular momentum and kinetic energy affected
    as she pulls her arms in toward her body as she spins?
    \begin{choices}
      \choice Her angular momentum remains constant, but her kinetic energy
      increases.
      \choice Her angular momentum increases, but her kinetic energy remains
      constant.
      \choice Her angular momentum decreases, but her kinetic energy remains
      constant.
      \choice Her angular momentum increases, but her kinetic energy decreases.
      \choice Both her angular momentum and kinetic energy remain constant.
    \end{choices}
    \vspace{.7in}
    
    \question A uniform rod of length $L$ and mass $m$ has a rotational inertia
    of $\dfrac1{12}mL^2$ about its center. A particle, also of mass $m$, is
    attached to one end of the stick. The combined rotational inertia of the
    stick and particle about the center of the rod is
    \begin{center}
      \begin{tikzpicture}[scale=3]
        \draw[thick](-1,-.03) rectangle (1,.03);
        \fill circle(.02) node[above]{$m$};
        \draw[thick,fill=gray!50](.85,-.1) rectangle(1,.1) node[midway]{$m$};
        \draw[|<->|](-1,-.2)--(1,-.2) node[midway,below]{$L$};
      \end{tikzpicture}
    \end{center}
    \begin{choices}
      \choice$\dfrac{mL^2}3$
      \choice$\dfrac{12mL^2}{13}$
      \choice$\dfrac{13mL^2}{12}$
      \choice$\dfrac{mL^2}{156}$
      \choice$\dfrac{13mL^2}{156}$
    \end{choices}
    \columnbreak
    
    \question Two disks are fixed to a vertical axle that is rotating with a
    constant angular speed $\omega$. The smaller disk has a mass $m$ and a
    radius $r$, and the larger disk has a mass $2m$ and radius $2r$. The
    general equation for the rotational inertia of a disk of mass $M$ and
    radius $R$ is $\dfrac12MR^2$. The ratio of the angular momentum of the
    larger disk to the smaller disk is
    \cpic{.25}{2disks}
    \begin{choices}
      \choice $1:4$
      \choice $4:1$
      \choice $1:2$
      \choice $2:1$
      \choice $8:1$
    \end{choices}

    \question A light rod has a mass attached at each end. At one end is a
    \SI6{\kilo\gram} mass, and at the other end is a \SI3{\kilo\gram} mass.
    An axis can be placed at any of the points shown. Through which point
    should an axis be placed so that the rotational inertia is the greatest
    about that axis?
    \begin{center}
      \begin{tikzpicture}[scale=1.5]
        \begin{scope}[thick]
          \draw circle(.1);
          \draw(4,0) circle(.15);
          \draw(.1,0)--(3.85,0);
        \end{scope}
        \fill (1,0) circle(.05);
        \fill (2,0) circle(.05);
        \fill (3,0) circle(.05);
        \node at (0,-.3){A};
        \node at (1,-.3){B};
        \node at (2,-.3){C};
        \node at (3,-.3){D};
        \node at (4,-.3){E};
        \node at (0,.3){3 kg};
        \node at (4,.3){6 kg};
      \end{tikzpicture}
    \end{center}
    \begin{choices}
      \choice A
      \choice B
      \choice C
      \choice D
      \choice E
    \end{choices}
    \columnbreak
    
    \uplevel{
      \textbf{Question \ref{lightrod1}--\ref{lightrod2}}:
      A light rod of negligible mass is pivoted at point $P$ a distance
      $L$ from one end as shown. A mass $m$ is attached to the left end of the
      rod at a distance of $3L$ from the pivot, and another mass $4m$ is
      attached to the other end a distance $L$ from the pivot. The system
      begins from rest in the horizontal position.
      \begin{center}
        \begin{tikzpicture}[scale=1.5]
          \draw[very thick](-3,0)--(1,0);
          \draw[thick](1,-.2) rectangle(1.4,.2) node[midway]{$4m$};
          \draw[thick](-3,-.11) rectangle(-3.22,.11) node[midway]{$m$};
          \fill circle(.05) node[above]{$P$};
          \node at (.5,.15){$L$};
          \node at (-1.5,.15){$3L$};
          \draw[thick,->,rotate=40](.35,0) arc(0:260:.35);
        \end{tikzpicture}
      \end{center}
    }
    
    \question The net torque acting on the system due to gravitational forces is
    \label{lightrod1}
    \begin{choices}
      \choice $4mgL$ clockwise
      \choice $3mgL$ clockwise
      \choice $3mgL$ counterclockwise
      \choice $mgL$ counterclockwise
      \choice $mgL$ clockwise
    \end{choices}
    \vspace{.6in}
    
    \question The angular acceleration of the system when it is released from
    rest is
    \label{lightrod2}
    \begin{choices}
      \choice zero
      \choice $\dfrac g{5L}$
      \choice $\dfrac g{4L}$
      \choice $\dfrac g{13L}$
      \choice $\dfrac gL$
    \end{choices}
    
    \question Two wheels are attached to each other and fixed so that they can
    only turn together. The smaller wheel has a radius of $r$ and the larger
    wheel has a radius of $3r$. The two wheels can rotate together on a
    frictionless axle. Three forces act tangentially on the edge of the wheels
    as shown. The magnitude of the net torque acting on the system of wheels is
    \begin{center}
      \begin{tikzpicture}[scale=.6]
        \begin{scope}[thick]
          \draw circle(1.2);
          \draw circle(3);
          \fill circle(.08);
        \end{scope}
        \begin{scope}[very thick]
          \draw[<->](0,0)--(0,-1.2) node[midway,right]{$r$};
          \draw[<->,rotate=-55](0,0)--(0,-3) node[midway,left]{$3r$};
          \draw[->](0,-1.2)--(2,-1.2) node[right]{$F$};
          \draw[->](1.2,0)--(1.2,2) node[left]{$F$};
          \draw[->](-3,0)--(-3,3) node[above]{$2F$};
        \end{scope}
      \end{tikzpicture}
    \end{center}
    \begin{choices}
      \choice $Fr$
      \choice $2Fr$
      \choice $3Fr$
      \choice $4Fr$
      \choice $6Fr$
    \end{choices}
  \end{questions}
  \setcounter{lastmc}{\value{question}}
\end{multicols*}
\newpage


%\genfreetitle{7}{ROTATIONAL MOTION, PART 1}{2}
%\genfreedirections

% TAKEN FROM THE 2015 AP PHYSICS C FREE-RESPONSE QUESTION MECH 3
\begin{center}
  \begin{tikzpicture}[scale=1.1]
    \draw[thick] rectangle(6,.28);
    \fill(.16,.14) circle(.08);
    \draw[|<->|](0,.55)--(6,.55) node[midway,fill=white]{$L$};
    \draw[<-](6,.14) to[in=270,out=0](6.5,.55) node[above]{$M$};
    \draw[<-](.15,.06) to[in=0,out=250](-.5,-.5) node[left]{Pivot Point};
  \end{tikzpicture}
\end{center}

\begin{questions}
  \setcounter{question}{\value{lastmc}}
  
  \question A uniform, thin rod of length $L$ and mass $M$ is allowed to pivot
  about its end, as shown in the figure above.
  \begin{parts}
    \part Using integral calculus, derive the rotational inertia for the rod
    around its end to show that it is $ML^2/3$.
    \vspace{\stretch1}
    
    \uplevel{
      \begin{center}
        \begin{tikzpicture}
          \draw[thick,dashed](0,.28) rectangle(.28,-5.72);
          \draw[thick,fill=white] rectangle(6,.28);
          \fill(.16,.14) circle(.08);
          \node[right] at (6,.14){$A$};
          \node[below] at (.14,-5.72){$B$};
          \draw[->,very thick](3,-.15) arc(-5:-85:2.8);
          \draw[thick,fill=white](.16,.14)--(.06,.7)--(.26,.7)--cycle;
          \fill[gray](-.4,.7) rectangle(.7,.9);
          \draw[thick](-.4,.7)--(.7,.7);
          %\draw[<-](.15,.06) to[in=0,out=250](-.5,-.5) node[left]{Pivot Point};
        \end{tikzpicture}
      \end{center}
      The rod is fixed at one end and allowed to fall from the horizontal
      position $A$ through the vertical position $B$.
    }
      
    \part Derive an expression for the velocity of the free end of the rod at
    position $B$. Express your answer in terms of $M$, $L$, and physical
    constants, as appropriate.
    \vspace{\stretch1}
    \newpage
    
    \uplevel{
      An experiment is designed to test the validity of the expression found
      in part (b). A student uses rods of various lengths that all have a
      uniform mass distribution. The student releases each of the rods from
      the horizontal position $A$ and uses photogates to measure the velocity
      of the free end at position $B$. The data are recorded below.
      \begin{center}
        \def\arraystretch{1.45}
        \begin{tabular}{|c|c|c|c|c|c|c|}
          \hline
          Length (m)     & 0.25 & 0.50 & 0.75 & 1.00 & 1.25 & 1.50\\\hline
          Velocity (m/s) & 2.7  & 3.8  & 4.6  & 5.2  & 5.8  & 6.3 \\\hline
          & & & & & & \\\hline
          & & & & & & \\\hline
        \end{tabular}
        \def\arraystretch{1}
      \end{center}
    }

    \part Indicate below which quantities should be graphed to yield a straight
    line whose slope could be used to calculate a numerical value for the
    acceleration due to gravity $g$.

    \vspace{.1in}Horizontal axis: \underline{\hspace{1in}}

    \vspace{.1in}Vertical axis: \underline{\hspace{1in}}

    \vspace{.1in}Use the remaining rows in the table above, as needed, to
    record any quantities that you indicated that are not given. Label each row
    you use and include units.

    \part Plot the straight line data points on the grid below. Clearly scale
    and label all axes, including units as appropriate. Draw a straight line
    that best represents the data.
    \begin{center}
      \begin{tikzpicture}[xscale=.58,yscale=.38]
        \draw[dashed,thick] grid(30,25);
        \draw[step=5,very thick] grid(30,25);
      \end{tikzpicture}
    \end{center}
    
    \part
    \begin{subparts}
      \subpart Using your straight line, determine an experimental value for
      $g$.
      \subpart Describe two ways in which the effects of air resistance could
      be reduced.
    \end{subparts}
  \end{parts}
  \newpage

  \uplevel{
    \cpic{.65}{disc1}
  }

  \question A large circular disk of mass $m$ and radius $R$ is initially
  stationary on a horizontal icy surface. A person of mass $m/2$ stands on the
  edge of the disk. Without slipping on the disk, the person throws a large
  stone of mass $m/20$ horizontally at initial speed $v_0$ from a height
  $h$ above the ice in a radial direction, as shown in the figures above. The
  coefficient of friction between the disk and the ice is $\mu$. All velocities
  are measured relative to the ground. The time it takes to throw the stone is
  negligible. Express all algebraic answers in terms of $m$, $R$, $v_0$,
  $h$, $m$, and fundamental constants, as appropriate.
  \begin{parts}
    \part Derive an expression for the length of time it will take the stone to
    strike the ice.
    \vspace{\stretch1}
    
    \part Assuming that the disk is free to slide on the ice, derive an
    expression for the speed of the disk and person immediately after the stone
    is thrown.
    \vspace{\stretch1}
    
    \part Derive an expression for the time it will take the disk to stop
    sliding.
    \vspace{\stretch1}
    \newpage
    
    \uplevel{
      \begin{center}
        \begin{tikzpicture}[scale=.85]
          \draw[very thick,fill=gray!70] circle(2);
          \draw[very thick,fill=black!60] circle(.1);
          \draw[very thick,fill=white](2,0) circle(.1);
          \fill (1.73,0) circle(.17);
          \draw[ultra thick,->](2,.1)--(2,1.5) node[above]{$v_0$};
        \end{tikzpicture}

        Top View
      \end{center}
      The person now stands on a similar disk of mass $m$ and radius $R$ that
      has a fixed pole through its center so that it can only rotate on the
      ice. The person throws the same stone horizontally in a tangential
      direction at initial speed $v_0$, as shown in the figure above. The
      rotational inertia of  the disk is $mR^2/2$.
    }

    \part Derive an expression for the angular speed $\omega$ of the disk
    immediately after the stone is thrown.\label{partd}
    \vspace{\stretch1}
    
    \part The person now stands on the disk at rest $R/2$ from the center of the
    disk. The person now throws the stone horizontally with a speed $v_0$ in
    the same direction as in part (\ref{partd}). Is the angular speed of the
    disk immediately after throwing the stone from this new position greater
    than, less than, or equal to the angular speed found in part (\ref{partd})?
    
    \vspace{.1in}
    \underline{\hspace{.3in}} Greater than\hspace{.2in}
    \underline{\hspace{.3in}} Less than\hspace{.2in}
    \underline{\hspace{.3in}} Equal to

    \vspace{.1in}Justify your answer.
    \vspace{\stretch1}
  \end{parts}
  \newpage
  
%  \question A uniform ball of radius $r$ rolls without slipping along the
%  loop-the-loop track in the figure below. The ball starts at rest at a height
%  of $h$ above the bottom of the loop.
%  \cpic{.55}{roll-ball}
%  \begin{parts}
%    \part If it is not to leave the track at the top of the loop, what is the
%    least value $h$ can have (in terms of radius $R$ of the loop)?
%
%    \part What would $h$ have to be if, instead of rolling, the ball slides
%    without friction?
%  \end{parts}
%  \newpage
%  
%  \begin{center}
%    \begin{tikzpicture}[scale=.65]
%      \fill[gray!70] circle(1);
%      \draw[ultra thick,->](-1,4.5)--(-1,0) arc(180:360:1)
%      --(1,2) node[above]{$F_A$};
%      \draw[thick](-1,0) arc(180:0:1);
%      \fill[gray](-5,4.5) rectangle(4,4.9);
%    \end{tikzpicture}
%  \end{center}
%  \question A disk of mass $M =\SI{2.}{\kilo\gram}$ and radius
%  $R=\SI{.10}\metre$ is supported by a rope of negligible mass, as shown
%  above. The rope is attached to the ceiling at one end and passes under the
%  disk. The other end of the rope is pulled upward with a force $F_A$. The
%  rotational inertia of the disk around its center is $MR^2/2$.
%  \begin{parts}
%    \part Calculate the magnitude of the force $F_A$ necessary to hold the disk
%    at rest.
%
%    \uplevel{
%      At time $t=0$, the force $F_A$ is increased to 12 N, causing the disk to
%      accelerate upward. The rope does not slip on the disk as the disk rotates.
%    }
%
%    \part Calculate the linear acceleration of the disk.
%    
%    \part Calculate the angular speed of the disk at $t=\SI{3.}\second$.
%    
%    \part Calculate the increase in total mechanical energy of the disk from
%    $t=0$ to $t=\SI{3.}\second$.
%    
%    \part The disk is replaced by a hoop of the same mass and radius. Indicate
%    whether the linear acceleration of the hoop is greater than, less than, or
%    the same as the linear acceleration of the disk. Justify your answer.
%
%    \vspace{.1in}   
%    \underline{\hspace{.3in}} Greater than\hspace{.2in}
%    \underline{\hspace{.3in}} Less than\hspace{.2in}
%    \underline{\hspace{.3in}} The same as
%  \end{parts}
\end{questions}
\end{document}
