\documentclass{../../../oss-classkick-exam}

\begin{document}
\genheader

\gentitle{C}{UNIVERSAL GRAVITATION}

\genmultidirections

\gengravity

\raggedcolumns
\begin{multicols*}{2}
  \begin{questions}
%    \question Two satellites of equal mass orbit a planet. Satellite B orbits
%    at twice the orbital radius of Satellite A. Which of the following
%    statements is true?
%    \begin{choices}
%      \choice The gravitational force on Satellite A is four times less than
%      that on Satellite B.
%      \choice The gravitational force on Satellite A is two times less than that
%      on Satellite B.
%      \choice The gravitational force on the satellites is equal.
%      \choice The gravitational force on Satellite A is two times greater than
%      that on Satellite B.
%      \choice The gravitational force on Satellite A is four times greater than
%      that on Satellite B.
%    \end{choices}
%    \vspace{.7in}
    
    \question A \SI{70}{\kilo\gram} astronaut floats at a distance of
    \SI{10}{\metre} from a \SI{50000}{\kilo\gram} spacecraft. What is the force
    of attraction between the astronaut and spacecraft?
    \begin{choices}
      \choice\SI{2.4e-6}{\newton}
      \choice\SI{2.4e-5}{\newton}
      \choice Zero; there is no gravity in space.
      \choice\SI{2.4e5}{\newton}
      \choice\SI{2.4e6}{\newton}
    \end{choices}
    \vspace{.5in}
    
%    \question The centripetal acceleration on \SI{1000}{\kilo\gram} car in a
%    turn is \SI{1e5}{\metre\per\second\squared}. The radius of the turn is
%    \SI{10}{\metre}. What is the car's speed?
%    \begin{choices}
%      \choice\SI{1e1}{\metre\per\second}
%      \choice\SI{1e2}{\metre\per\second}
%      \choice\SI{1e3}{\metre\per\second}
%      \choice\SI{1e4}{\metre\per\second}
%      \choice\SI{1e5}{\metre\per\second}
%    \end{choices}
    
    \question The Earth is at an average distance of \SI{1}{AU} from the Sun
    and has an orbital period of \SI{1}{year}. Jupiter orbits the Sun at
    approximately \SI{5}{AU}. About how long is the orbital period of Jupiter?
    \begin{choices}
      \choice\SI{1}{year}
      \choice\SI{2}{years}
      \choice\SI{5}{years}
      \choice\SI{11}{years}
      \choice\SI{125}{years}
    \end{choices}
    
    \question A proposed ``space elevator'' can lift a \SI{1000}{\kilo\gram}
    payload to an orbit of \SI{150}{\kilo\metre} above the Earth's surface. The
    radius of the Earth is \SI{6.4e6}{\metre}, and the Earth's mass is
    \SI{6.e24}{\kilo\gram}. What is the gravitational potential energy of the
    payload when it reaches orbit?
    \begin{choices}
      \choice\SI{1.0e3}{\joule}
      \choice\SI{2.7e6}{\joule}
      \choice\SI{6.1e10}{\joule}
      \choice\SI{2.7e12}{\joule}
      \choice\SI{1.0e15}{\joule}
    \end{choices}
    
    \question A satellite orbits the Earth at a distance of
    \SI{200}{\kilo\metre}. If the mass of the Earth is \SI{6.e24}{\kilo\gram}
    and the Earth's radius is \SI{6.4e6}{\metre}, what is the satellite's speed?
    \begin{choices}
      \choice\SI{1.e3}{\metre\per\second}
      \choice\SI{3.5e3}{\metre\per\second}
      \choice\SI{7.8e3}{\metre\per\second}
      \choice\SI{5e6}{\metre\per\second}
      \choice\SI{6.1e7}{\metre\per\second}
    \end{choices}
    
    \question Mars orbits the Sun at a distance of \SI{2.3e11}{\metre}. The
    mass of the Sun is \SI{2.e30}{\kilo\gram}, and the mass of Mars is
    \SI{6.4e23}{\kilo\gram}. Approximately what is the gravitational force that
    the Sun exerts on Mars?
    \begin{choices}
      \choice\SI{1.6e20}{\newton}
      \choice\SI{1.6e21}{\newton}
      \choice\SI{3.7e21}{\newton}
      \choice\SI{3.7e32}{\newton}
      \choice\SI{3.7e42}{\newton}
    \end{choices}

    \question The mass of a planet is $1/4$ that of Earth and its radius is
    half of Earth's radius. The acceleration due to gravity on this planet is
    most nearly
    \begin{choices}
      \choice\SI{2}{\metre\per\second\squared}
      \choice\SI{4}{\metre\per\second\squared}
      \choice\SI{5}{\metre\per\second\squared}
      \choice\SI{10}{\metre\per\second\squared}
      \choice\SI{20}{\metre\per\second\squared}
    \end{choices}
    
    \question When climbing from sea level to the top of Mount Everest, a hiker
    changes elevation by \SI{8848}{\metre}. By what percentage will the
    gravitational field of the Earth change during the climb? (The Earth's
    mass is \SI{6.e24}{\kilo\gram}, and its radius is \SI{6.4e6}{\metre}.)
    \begin{choices}
      \choice It will increase by approximately \SI{.3}{\percent}.
      \choice It will decrease by approximately \SI{.3}{\percent}.
      \choice It will increase by approximately \SI{12}{\percent}.
      \choice It will decrease by approximately \SI{12}{\percent}.
      \choice The gravitational field strength will not change.
    \end{choices}
    \vspace{.7in}
    
    \question Four planets, A through D, orbit the same star. The relative
    masses and distances from the star for each planet are shown in the table.
    For example, Planet A has twice the mass of Planet B, and Planet D has
    three times the orbital radius of Planet A. Which planet has the highest
    gravitational attraction to the star?
    \begin{center}
      \vspace{-.1in}
      \begin{tabular}{lll}
        \hline
        \textbf{Planet} & \textbf{Relative mass} & \textbf{Relative distance}\\
        \hline
        A\hspace{.4in}& $2m$     & $r$    \\ \hline
        B & $m$                  & $0.1r$\hspace{.25in} \\ \hline
        C & $0.5m$\hspace{.25in} & $2r$   \\ \hline
        D & $4m$                 & $3r$   \\ \hline
      \end{tabular}
    \end{center}
    \begin{choices}
      \choice Planet A
      \choice Planet B
      \choice Planet C
      \choice Planet D
      \choice All have the same gravitational attraction to the star.
    \end{choices}
    \vspace{.7in}
    
    \question A satellite orbits the Earth at a distance that is four times the
    radius of the Earth. If the acceleration due to gravity near the surface of
    the Earth is $g$, the acceleration of the satellite is most nearly
    \begin{choices}
      \choice zero
      \choice $\dfrac{g}2$
      \choice $\dfrac{g}4$
      \choice $\dfrac{g}8$
      \choice $\dfrac{g}{16}$
    \end{choices}
  
    \question A satellite orbits the Earth in an elliptical orbit, with point A
    being close to the Earth and point B farther away. As the satellite moves
    from point A to point B, which of the following is true of the angular
    momentum and kinetic energy of the satellite?
    \begin{center}
      \vspace{-.1in}
      \begin{tikzpicture}[scale=1.4]
        \tikzstyle{balloon}=[ball color=gray];
        \draw(0,0) ellipse (2 and 1);
        \draw[fill=black](2,0) circle(0.05) node[right]{A};
        \draw[fill=black](-2,0) circle(0.05) node[left]{B};
        \shade[balloon] (0.5,0) circle (0.2);
      \end{tikzpicture}
    \end{center}
  
    \begin{tabular}{lll}
      & \underline{Angular momentum} & \underline{Kinetic energy}\\
      (A) & Increases & Remains constant \\
      (B) & Remains constant & Increases \\
      (C) & Decreases & Remains constant \\
      (D) & Remains constant & Decreases \\
      (E) & Remains constant & Remains constant
    \end{tabular}
    \vspace{.7in}
    
    \question Two planets of mass $M$ and $9M$ are in the same solar system. The
    radius of the planet of mass $M$ is $R$. In order for the acceleration due
    to gravity to be the same for each planet, the radius of the planet of mass
    $9M$ would have to be
    \begin{choices}
      \choice $\dfrac{R}2$
      \choice $R$
      \choice $2R$
      \choice $3R$
      \choice $9R$
    \end{choices}
    
    \question A satellite is in a stable circular orbit around the Earth at a
    radius $R$ and speed $\varv$. At what radius would the satellite travel in
    a stable orbit with a speed $2\varv$?
    \begin{choices}
      \choice $\dfrac{R}4$
      \choice $\dfrac{R}2$
      \choice $R$
      \choice $2R$
      \choice $4R$
    \end{choices}

%    \question Two masses exert a gravitational force $F$ on each other. If one
%    of the masses is doubled, and the distance between the masses is tripled,
%    the new force between them is
%    \begin{choices}
%      \choice $6F$
%      \choice $\dfrac{2F}3$
%      \choice $\dfrac{2F}9$
%      \choice $\dfrac{3F}2$
%      \choice $\dfrac{4F}9$
%    \end{choices}
    
    \question Two planets, X and Y, orbit a star. Planet X orbits at a radius
    $R$, and Planet Y orbits at a radius $3R$. Which of the following best
    represents the relationship between the acceleration $a_X$ of Planet X and
    the acceleration $a_Y$ of Planet Y?
    \begin{center}
      \begin{tikzpicture}[scale=.7]
        \tikzstyle{balloon}=[ball color=gray];
        \shade[balloon](0,0) circle(.25);
        \draw[dashed](0,0) circle(1);
        \draw[dashed](0,0) circle(3);
        \draw(0,0)--(1,0) node[right]{X} node[midway,above]{$R$};
        \draw[fill=black](1,0) circle(.05);
        \begin{scope}[rotate=230]
          \draw(0,0)--(3,0) node[left]{Y} node[pos=0.7,right]{$3R$};
          \draw[fill=black](3,0) circle(.05);
        \end{scope}
      \end{tikzpicture}
    \end{center}
    \begin{choices}
      \choice $a_X = 9a_Y$
      \choice $9a_X = a_Y$
      \choice $a_X = 3a_Y$
      \choice $3a_X = a_Y$
      \choice $a_X = a_Y$
    \end{choices}

%    \question The Earth and the moon apply a gravitational force to each other.
%    Which of the following statements is true?
%    \begin{choices}
%      \choice The Earth applies a greater force on the moon than the moon
%      exerts on the Earth.
%      \choice The Earth applies a smaller force on the moon than the moon
%      exerts on the Earth.
%      \choice The Earth applies a force on the moon, but the moon does not
%      exert a force on the Earth.
%      \choice The Earth does not apply a force on the moon, but the moon exerts
%      a force on the Earth.
%      \choice The force the Earth applies to the moon is equal and opposite to
%      the force the moon applies to the Earth.
%    \end{choices}
%    \vspace{.7in}
    
    \question A planet orbits at a radius $R$ around a star of mass $M$. The
    period of orbit of the planet is
    \begin{choices}
      \choice $\sqrt{\dfrac{4\pi^2R^2}{GM}}$
      \choice $\dfrac{4\pi^2R^3}{GM}$
      \choice $\sqrt{\dfrac{4\pi^2R^3}{GM}}$
      \choice $\sqrt{\dfrac{4\pi^2R}{GM}}$
      \choice $\dfrac{GM}{4\pi^2R}$
    \end{choices}

    \question A moon orbits a large planet in an elliptical orbit, with its
    closest approach at a distance $a$, and its farthest distance $b$. The
    speed of the moon at point b is $\varv$. The speed at point $a$ is
    \begin{choices}
      \choice $\dfrac{a\varv}{b}$
      \choice $\dfrac{b\varv}{a}$
      \choice $\dfrac{(a+b)\varv}{b}$
      \choice $\dfrac{(b-a)\varv}{b}$
      \choice $\dfrac{2b\varv}{a}$
    \end{choices}

    \question A satellite orbits the Earth in an elliptical orbit. Which of the
    following statements is true?
    \begin{choices}
      \choice The angular velocity of the satellite increases as it travels
      farther from the Earth.
      \choice The acceleration of the satellite increases as it travels closer
      to the Earth.
      \choice The angular momentum of the satellite increases as it travels
      closer to the Earth.
      \choice The potential energy of the satellite is equal to its kinetic
      energy at all points in the orbit.
      \choice The speed of the satellite must remain constant for it to remain
      in orbit around the Earth.
    \end{choices}
    \vspace{.7in}
    
%    \question If a planet has twice the radius of Earth and half of Earth's
%    density, what is the acceleration due to gravity on the surface of the
%    planet, in terms of the gravitational acceleration $g$ on the surface of
%    Earth?
%    \begin{choices}
%      \choice $4g$
%      \choice $2g$
%      \choice $g$
%      \choice $\dfrac{g}2$
%      \choice $\dfrac{g}4$
%    \end{choices}
    \columnbreak

    \uplevel{
      \textbf{Questions \ref{moon1}--\ref{moon2}}
      
      Two moons of mass $m$ and $2m$ orbit a planet of mass $M$ at the same
      radius $R$ and speed $\varv$ toward each other, as shown. The moons
      collide and stick together without destroying either moon.
      \begin{center}
        \begin{tikzpicture}[scale=.8]
          \tikzstyle{balloon}=[ball color=gray];
          \shade[balloon] circle(.25) node[below]{$M$};
          \draw[dashed](0,0) circle(3);
          \begin{scope}[rotate=50]
            \draw[->](0.25,0)--(2.8,0) node[midway,right]{$R$};
            \shade[balloon] (3,0) circle (0.2) node[right]{$m$};
            \draw[thick,->](3,0.25)--(3,1.25) node[above]{$\varv$};
          \end{scope}
          \begin{scope}[rotate=130]
            \shade[balloon] (3,0) circle (0.2) node[left]{$2m$};
            \draw[thick,->](3,-0.25)--(3,-1.25) node[above]{$\varv$};
          \end{scope}
        \end{tikzpicture}
      \end{center}
    }
    
    \question The total momentum of the moons after the collision is
    \label{moon1}
    \begin{choices}
      \choice $m\varv$
      \choice $2m\varv$
      \choice $3m\varv$
      \choice $6m\varv$
      \choice zero
    \end{choices}
    
    \question The velocity of the two masses after the collision above is
    \label{moon2}
    \begin{choices}
      \choice $\varv$ counterclockwise
      \choice $\varv/2$ counterclockwise
      \choice $\varv/2$ clockwise
      \choice $\varv/3$ counterclockwise
      \choice $\varv/3$ clockwise
    \end{choices}
    \vspace{.7in}
    
%  \question Consider a two-star system shown above, which consists of two stars of
%    mass $m$ rotating in a circle of radius $r$ about their center of mass. What
%    is the total energy of the two-star system?
%    \begin{choices}
%    \choice $-Gm^2/2r$
%    \choice $Gm^2/2r$
%    \choice $Gm^2/4r$
%    \choice $3Gm^2/4r$
%    \choice $-Gm^2/4r$
%    \end{choices}
    
    \question A satellite of mass $m$ travels in an elliptical orbit around a
    planet of mass $M$. The satellite has a speed $\varv$ when it is closest to
    the planet at a distance $r$. Work is done by the engines of the satellite
    to change its orbit to a circular orbit when it is at this distance $r$.
    Which of the following statements is true of the transition from an
    elliptical orbit to a circular orbit?
    \begin{choices}
      \choice The work done by the satellite engines to change the orbit is
      equal to the change in kinetic energy of the satellite.
      \choice The work done by the satellite engines to change the orbit is
      equal to the change in potential energy of the satellite.
      \choice The work done by the satellite engines to change the orbit is
      equal to the change in angular momentum of the satellite.
      \choice The work done by the satellite engines to change the orbit is
      equal to the change in speed of the satellite.
      \choice The work done by the satellite engines to change the orbit is
      equal to the change in orbital radius of the satellite.
    \end{choices}
    \vspace{.7in}
    
    \question A satellite of mass $m$ orbits the Earth with a potential energy
    $U$ and a kinetic energy $K$. Which of the following statements would have
    to be true for the satellite to escape the Earth's gravity completely?
    \begin{choices}
      \choice The kinetic energy of the satellite would have to be equal to the
      potential energy between the Earth and the satellite.
      \choice The potential energy between the Earth and the satellite would
      have to be greater than the kinetic energy of the satellite.
      \choice The total energy of the satellite would have to be greater than
      the kinetic energy of the satellite.
      \choice The kinetic energy of the satellite would have to be greater than
      the potential energy of the satellite.
      \choice The total energy of the satellite would have to be equal to the
      potential energy of the satellite.
    \end{choices}
  \end{questions}
\end{multicols*}
\newpage

\genfreetitle{C}{UNIVERSAL GRAVITATION}{4}

\genfreedirections

\begin{questions}

%\item A spacecraft moving with an initial velocity $\mb{v}_0$, shown below,
%  ``slingshots'' around the sun in order to reverse its direction. The sun's
%  mass is $m_\mathrm{sun}$ and you can make the assumption that the sun remains
%  stationary.\\
%  \begin{minipage}{0.28\textwidth}
%    \pic{1}{shuttle.jpg}
%  \end{minipage}
%  \begin{minipage}{0.7\textwidth}
%    \begin{enumerate}[noitemsep]
%    \item What is the minimum initial speed required by the spacecraft to escape
%      the sun's gravitational field and move in a trajectory toward infinity?
%    \item What is the minimum initial speed $v_o$ that the spacecraft must have
%      in order to avoid falling into the sun? (Treat the sun and the spacecraft
%      as points.)
%    \item Repeat the previous questions, but now the sun has a radius $R$.
%    \item Write down the equations required to calculate the initial angle
%      $\theta$ in terms of $v_0$, $d$, $m_\mathrm{sun}$, $G$, and $r$.
%    \end{enumerate}
%  \end{minipage}
%  \newpage

  % TAKEN FROM THE 2007 AP PHYSICS C MECHANICS EXAM FREE-RESPONSE QUESTION 2
  \question In March 1999 the Mars Global Surveyor (GS) entered its final orbit
  about Mars, sending data back to Earth. Assume a circular orbit with a period
  of $\SI{1.18e2}{minutes}=\SI{7.08e3}{\second}$ and orbital speed of
  \SI{3.40e3}{\metre\per\second}. The mass of the GS is \SI{930}{\kilo\gram}
  and the radius of Mars is \SI{3.43e6}{\metre}.
  \begin{parts}
    \part Calculate the radius of the GS orbit.
    \part Calculate the mass of Mars.
    \part Calculate the total mechanical energy of the GS in this orbit.
    \part If the GS was to be placed in a lower circular orbit (closer to the
    surface of Mars), would the new orbital period of the GS be greater than or
    less than the given period? Justify your answer.

    \vspace{.15in}
    \underline{\hspace{.3in}} Greater than\hspace{1in}
    \underline{\hspace{.3in}} Less than

    \part In fact, the orbit the GS entered was slightly elliptical with its
    closest approach to Mars at \SI{3.71e5}{\metre} above the surface and its
    furthest distance at \SI{4.36e5}{\metre} above the surface. If the speed of
    the GS at closest approach is \SI{3.40e3}{\metre\per\second}, calculate the
    speed at the furthest point of the orbit.
  \end{parts}
  \newpage
  
  % TAKEN FROM THE 2005 AP PHYSICS C MECHANICS EXAM FREE-RESPONSE QUESTION 2
  \question A student is given the set of orbital data for some of the moons of
  Saturn shown below and is asked to use the data to determine the mass $M_S$
  of Saturn. Assume the orbits of these moons are circular.
  \begin{center}
    \def\arraystretch{1.5}
    \begin{tabular}{|c|c|p{1in}|p{1in}|}
      \hline
      Orbital Period, $T$ &  Orbital Radius, $R$ & & \\
      (seconds)           &  (meters)            & & \\
      \hline
      \num{8.14e4} & \num{1.85e8} & & \\ \hline
      \num{1.18e5} & \num{2.38e8} & & \\ \hline
      \num{1.63e5} & \num{2.95e8} & & \\ \hline
      \num{2.37e5} & \num{3.77e8} & & \\ \hline
    \end{tabular}
  \end{center}
  \def\arraystretch{1}
  \begin{parts}
    \part Write an algebraic expression for the gravitational force between
    Saturn and one of its moons.
    \label{algebraic}
    
    \part Use your expression from part (\ref{algebraic}) and the assumption of
    circular orbits to derive an equation for the orbital period $T$ of a moon
    as a function of its orbital radius $R$.
    
    \part Which quantities should be graphed to yield a straight line whose
    slope could be used to determine Saturn's mass?

    \part Complete the data table by calculating the two quantities to be
    graphed. Label the top of each column, including units.

    \part Plot the graph on the axes below. Label the axes with the variables
    used and appropriate numbers to indicate the scale.
    \cpic{.8}{graph-paper}
    \part Using the graph, calculate a value for the mass of Saturn.
  \end{parts}
  \newpage
  
%\item Two stars of equal mass $M$ are orbiting each other in a circular path.
%  Show that the orbital period is given by:
%  \begin{displaymath}
%    T^2=\frac{2\pi^2d^3}{GM}
%  \end{displaymath}
%  where $d$ is the distance between the stars.
%  \newpage
  
  \question A planet of mass $M$, radius $R$, and uniform density has a small
  tunnel drilled through the center of the planet, as shown below. When the
  mass is inside the tunnel, it experiences a force of $F=(GmM/R^3)r$, whereas
  when the mass is outside of the planet, it experiences a gravitational force
  of $F=GmM/r^2$.
  \begin{center}
    \begin{tikzpicture}
      \draw(0,0) circle(2);
      \draw(-0.1,2) rectangle(0.1,-2);
      \draw[->,thick,rotate=35](0,0)--(2,0) node[midway,below]{$R$};
    \end{tikzpicture}
  \end{center}
  \begin{parts}
    \part Setting the potential energy of the mass to be zero at the planet's
    center, calculate the mass's potential energy as a function of distance from
    the center of the planet $U(r)$, for values $r<R$. Sketch this potential
    function.
    
    \part If the mass is dropped from $R$ from the center of the planet, how
    long will it take until it returns to its original position?

    \part If the mass is dropped from $R/2$ from the center of the planet, will
    it require more, or less, or the same amount of time to return to its
    original position compared to if it was dropped from $R$?

    \part If the mass is dropped from $2R$ from the center of the planet, will
    it require more, or less, or the same amount of time to return to its
    original position compared to if it was dropped from $R$?
  \end{parts}
  \newpage

  \question Two stars of unequal mass orbit each other about their common
  center of mass as shown. The star of mass $M_1$ orbits in a circle of radius
  $r$, and the star of mass $M_2$ orbits in a circle of radius $2r$.
  \begin{center}
    \begin{tikzpicture}[scale=.8]
      \tikzstyle{balloon}=[ball color=gray];
      \draw[fill=black](0,0) circle(0.075);
      \draw[dashed](0,0) circle(3);
      \draw[dashed](0,0) circle(1.5);
      \draw[->](0,0)--(1.5,0) node[midway,below]{$r$};
      \shade[balloon] (1.5,0) circle (0.2) node[right]{$M_1$};
      \draw[->](0,0)--(-3,0) node[midway,below]{$2r$};
      \shade[balloon] (-3,0) circle (0.2) node[left]{$M_2$};
    \end{tikzpicture}
  \end{center}
  \begin{parts}
    \part Determine the ratio of masses $M_1/M_2$.
    \part Determine the ratio of the acceleration $a_1$ of $M_1$ to the
    acceleration $a_2$ of $M_2$.
    \part Determine the ratio of the period $T_1$ of $M _1$ to the period $T_2$
    of $M_2$.
  \end{parts}
  \newpage
  

%\item Five equal masses $M$ are equally spaced on the arc of a semicircle of
%  radius $R$ as shown in the figure below. A mass $m$ is located at the center
%  of curvature of the arc. If $M$ is \SI{3}{\kilo\gram}, $m$ is
%  \SI{21}{\kilo\gram}, nad $R$ is \SI{10}{\centi\metre}, what is the force on
%  $m$ due to the five masses?
%  \begin{center}
%    \begin{tikzpicture}[scale=.6]
%      \tikzstyle{balloon1}=[ball color=gray!50];
%      \tikzstyle{balloon2}=[ball color=red!60!black];
%      \draw[->](0,0)--(5,0) node[right]{$x$};
%      \draw[->](0,0)--(0,5) node[above]{$y$};
%      \draw[ultra thick,blue!80](4,0) arc(0:180:4);
%      \shade[balloon1] (0,0) circle (.1) node[below]{$m$};
%      \foreach \theta in {0,45,...,180} {
%        \begin{scope}[rotate=\theta]
%          \shade[balloon2] (4,0) circle (.3) node[white]{$M$};
%        \end{scope}
%      }
%    \end{tikzpicture}
%  \end{center}
%  \vspace{\stretch{1}}
%  \newpage

  \question Two point particles of mass $m$ are on the $y$ axis at $y=a$ and
  $y=-a$, as shown in the figure below.
  \begin{center}
    \begin{tikzpicture}[scale=.7]
      \tikzstyle{balloon1}=[ball color=green!50!gray];
      \tikzstyle{balloon2}=[ball color=red!50];
      \draw[->](-2,0)--(6,0) node[right]{$x$};
      \draw[->](0,-3)--(0,3) node[above]{$y$};
      \draw[<->](-1,0)--(-1,2) node[midway,left]{$a$};
      \draw[<->](-1,0)--(-1,-2)node[midway,left]{$a$};
      \draw(0,2)--(-1.2,2);
      \draw(0,-2)--(-1.2,-2);
      \shade[balloon1] (0,2) circle (.55) node[white]{$m$};
      \shade[balloon1] (0,-2)circle (.55) node[white]{$m$};
      \shade[balloon2] (4,0) circle (.45) node[white]{$m_0$};
    \end{tikzpicture}
  \end{center}
  \begin{parts}
    \part Derive the expression for the gravitational force exerted by these two
    particles on a third particle of mass $m_0$ located on the $x$ axis at a
    distance $x$ away from the origin.
    
    \part What is the gravitational field $\mb{g}$ on the $x$-axis due to the
    two particles?
    
    \part Show that $g_x$ (the $x$ component of $\mb{g}$) due to the two
    particles on the $y$ axis is approximately $\displaystyle-\frac{2Gm}{x^2}$
    when $x$ is much greater than $a$.
    
    \part Show that the maximum value of $|g_x|$ occurs at the point
    $\displaystyle x=\frac{\pm a}{\sqrt{2}}$.
  \end{parts}
  \newpage

  
%\item\textbf{THIS IS A CHALLENGE PROBLEM THAT IS MORE DIFFICULT THAN AP EXAMS:}
%  Spacecraft that study the Sun are often placed at the ``L1 Lagrange Point'',
%  located sunward of Earth on the Sun--Earth line. L1 is the point where Earth's
%  and Sun's gravity together produce an orbital period of one year, so that a
%  spacecraft at L1 stays fixed relative to Earth as both planet and spacecraft
%  orbit the Sun. This placement ensures an uninterrupted view of the sun,
%  without being periodically eclipsed by Earth as would occur in Earth orbit.
%  Find L1's location relative to Earth. (Hint: This problem calls for numerical
%  methods for solving high-order polynomial equation.)

\end{questions}
\end{document}
