\documentclass{../../../oss-classkick}

\begin{document}
\genheader

\gentitle{2}{FLUID MECHANICS}

\genmultidirections

\gengravity

\raggedcolumns
\begin{multicols}{2}

  \begin{enumerate}[leftmargin=18pt]

  \item Two blocks of different sizes and masses float in a tray of water. Each
    block is half submerged, as shown in the figure. Water has a density of
    \SI{1000}{\kilo\gram\per\metre\cubed}. What can be concluded about the
    densities of the two blocks?
    
    \pic{.39}{mc-q1}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\vspace{-.1in} The two blocks have different densities, both of which
      are less than \SI{1000}{\kilo\gram\per\metre\cubed}.
    \item The two blocks have the same density of
      \SI{500}{\kilo\gram\per\metre\cubed}.
    \item The two blocks have the same density, but the density cannot be
      determined with the information given.
    \item The larger block has a greater density than the smaller block, but
      the densities of the blocks cannot be determined with the information
      given.
    \end{enumerate}
    \vspace{.65in}
    
  \item The figure shows four cylinders of various diameters filled to different
    heights with water. A hole in the side of each cylinder is plugged by a
    cork. All cylinders are open at the top. The corks are removed. Which
    of the following is the correct ranking of the velocity of the water ($v$)
    as it exits each cylinder?
    
    \pic{.4}{mc-q2}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $v_A > v_D > v_C > v_B$
    \item $v_A = v_D > v_C > v_B$
    \item $v_B > v_C > v_A = v_D$
    \item $v_C > v_A = v_B = v_D$
    \end{enumerate}
    \vspace{.65in}

  \item A \SI1{\centi\metre} diameter pipe leads to a showerhead with twenty
    \SI{1}{\milli\metre} diameter exit holes. The velocity of the water in the
    pipe is $v$. What is the velocity of the water exiting the holes?
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $0.05v$
    \item $0.5v$
    \item $5v$
    \item $100v$
    \end{enumerate}
  \end{enumerate}
  \columnbreak
    
  \textbf{Questions \ref{cyl1} and \ref{cyl2}}

  Four differently shaped sealed containers are completely filled with alcohol,
  as shown below. Containers $A$ and $B$ are cylindrical. Containers $C$ and
  $D$ are truncated conical shapes. The top and bottom diameters of the
  containers are shown.
  \pic{.45}{mc-q3-4}
  \begin{enumerate}[leftmargin=18pt,resume]
    
  \item Which of the following is the correct ranking of the pressure ($P$) at
    the bottom of the containers?
    \label{cyl1}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $P_A = P_B = P_C = P_D$
    \item $P_A = P_D > P_C = P_B$
    \item $P_A > P_D > P_C > P_B$
    \item $P_D > P_A > P_C > P_B$
    \end{enumerate}
    \vspace{.7in}
    
  \item The force on the bottom of container $A$ due to the fluid inside the
    container is $F$. What is the force on the bottom of container $B$ due to
    the fluid inside?
    \label{cyl2}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $F$
    \item $F/4$
    \item $F/8$
    \item $F/16$
    \end{enumerate}

  \item Two cylinders filled with a fluid are connected by a pipe so that fluid
    can pass between the cylinders, as shown in the figure. The cylinder on the
    right has 4 times the diameter of the cylinder on the left. Both cylinders
    are fitted with a movable piston and a platform on top. A person stands on
    the left platform. Which of the following lists the correct number of
    people that need to stand on the right platform so neither platform moves.
    Assume that the platform and piston have negligible mass and that all the
    people have the same mass.

    \vspace{-.1in}\pic{.35}{mc-q5}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item \num{16} people
    \item \num{4} people
    \item \num{1} person
    \item It is impossible to balance the system because you need $1/16$ of a
      person on the right side.
    \end{enumerate}
    \vspace{.7in}
    \columnbreak
    
  \item A mass $m$ is suspended in a fluid of density $\rho$ by a string, as
    shown in the figure below. The tension in the string is $T$. Which of the
    following is an appropriate equation for the buoyancy force? Select two
    answers.
    \begin{center}
      \vspace{-.15in}\pic{.3}{mc-q6}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $F_b=mg$
    \item $F_b=mg-T$
    \item $F_c=a_2 \rho gh_1$
    \item $F_d=a\rho g(h_1-h_2)$
    \end{enumerate}
    \vspace{.7in}
    
  \item Three wooden blocks of different masses and sizes float in a container
    of water, as shown in the figure. Each of the masses has a weight on top.
    Which of the following correctly ranks the buoyancy force on the wooden
    blocks?
    \begin{center}
      \vspace{-.15in}\pic{.35}{mc-q7}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $A > B = C$
    \item $A = B > C$
    \item $B > A = C$
    \item $B > A > C$
    \end{enumerate}
    
  \item Two blocks of the same dimensions are floating in a container of water,
    as shown in the figure. Which of the following is a correct statement about
    the two blocks?
    \begin{center}
      \vspace{-.15in}\pic{.35}{mc-q8}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item The net force on both blocks is the same.
    \item The buoyancy force exerted on both blocks is the same.
    \item The density of both blocks is the same.
    \item The pressure exerted on the bottom of each block is the same.
    \end{enumerate}
    \vspace{.7in}
    
  \item The figure shows four cubes of the same volume at rest in a container
    of water. Cube $C$ is partially submerged. Cubes $A$, $B$, and $D$ are fully
    submerged, with $B$ resting on the bottom of the container. Which of the
    following correctly ranks the densities ($\rho$) of the cubes? Assume the
    water to be incompressible.
    \begin{center}
      \vspace{-.15in}
      \pic{.3}{mc-q9}
    \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item $\rho_C >\rho_D >\rho_A >\rho_B$
    \item $\rho_B >\rho_A >\rho_D >\rho_C$
    \item $\rho_B >\rho_A =\rho_D >\rho_C$
    \item $\rho_B >\rho_A =\rho_D =\rho_C$
    \end{enumerate}
    \columnbreak
    
  \item A beaker of water sits on a balance. A metal block with a mass of
    \SI{70}{\gram} is held suspended in the water by a spring scale in position
    1, as shown. In this position, the reading on the balance is
    \SI{1260}{\gram}, and the spring scale reads \SI{120}{\gram}. When the
    block is lifted from the water to position 2, what are the readings on the
    balance and spring scale?
    \begin{center}
      \vspace{-.15in}
      \pic{.2}{mc-q10}

      \begin{tabular}{c c c}
        & \textbf{Balance reading} & \textbf{Spring scale reading}\\
        (A) & \SI{1190}{\gram} & \SI{120}{\gram}\\
        (B) & \SI{1190}{\gram} & \SI{190}{\gram}\\
        (C) & \SI{1260}{\gram} & \SI{120}{\gram}\\
        (D) & \SI{1330}{\gram} & \SI{120}{\gram}
      \end{tabular}
    \end{center}
    
  \item Blood cells pass through an artery that has a buildup of plaque along
    both walls, as shown in the figure. Which of the following correctly
    describes the behavior of the blood cells as they move from the right
    side of the figure through the area of plaque? Assume the blood cells
    can change volume.
    \begin{center}
      \vspace{-.15in}
      \pic{.25}{mc-q11}
     \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\vspace{-.1in} The blood cells increase in speed and expand in volume.
    \item The blood cells increase in speed and decrease in volume.
    \item The blood cells decrease in speed and expand in volume.
    \item The blood cells decrease in speed and decrease in volume.
    \end{enumerate}
    \vspace{.7in}
    
  \item Firefighters use a hose with a 2 cm exit nozzle connected to a hydrant
    with an 8 cm diameter opening to attack a fire on the second floor of a
    building 6 m above the hydrant, as shown in the figure. What pressure must
    be supplied at the hydrant to produce an exit velocity of
    \SI{15}{\metre\per\second}? (The density of water is
    \SI{1000}{\kilo\gram\per\metre\cubed}, and the exit pressure is
    \SI{1e5}{\pascal}.)
    \begin{center}
      \vspace{-.2in}
      \pic{.4}{mc-q12}
     \end{center}
    \begin{enumerate}[nosep,leftmargin=18pt,label=(\Alph*)]
    \item\SI{1.7e5}\pascal
    \item\SI{2.0e5}\pascal
    \item\SI{2.6e5}\pascal
    \item\SI{3.2e5}\pascal
    \end{enumerate}
  \end{enumerate}
\end{multicols}

%\genanswersheet{2}{Fluid Mechanics \& Thermodynamics}
%
%\begin{center}
%  \begin{minipage}[t]{.2\textwidth}
%    \vspace{.2in}
%    \bgroup
%    \begin{tabular}{>{\centering}m{1.3cm} >{\centering}m{1.7cm}}
%      No. & Answer
%    \end{tabular}\\
%    \def\arraystretch{1.5}
%    \begin{tabular}{|>{\centering}m{1.3cm}|>{\centering}m{1.7cm}|}
%      \hline
%      1 & \\ \hline
%      2 & \\ \hline
%      3 & \\ \hline
%      4 & \\ \hline
%      5 & \\ \hline
%      6 & \\ \hline
%      7 & \\ \hline
%      8 & \\ \hline
%      9 & \\ \hline
%      10 & \\ \hline
%      11 & \\ \hline
%      12 & \\ \hline
%      13 & \\ \hline
%      14 & \\ \hline
%      15 & \\ \hline
%      16 & \\ \hline
%      17 & \\ \hline
%      18 & \\ \hline
%      19 & \\ \hline
%      20 & \\ \hline
%      21 & \\ \hline
%      22 & \\ \hline
%      23 & \\ \hline
%      24 & \\ \hline
%      25 & \\ \hline
%    \end{tabular}
%    \egroup
%  \end{minipage}
%  \hspace{.5in}
%  \begin{minipage}[t]{.2\textwidth}
%    \vspace{.2in}
%    \bgroup
%    \begin{tabular}{>{\centering}m{1.3cm} >{\centering}m{1.7cm}}
%      No. & Answer
%    \end{tabular}\\
%    \def\arraystretch{1.5}
%    \begin{tabular}{|>{\centering}m{1.3cm}|>{\centering}m{1.7cm}|}
%      \hline
%      26 & \\ \hline
%      27 & \\ \hline
%      28 & \\ \hline
%      29 & \\ \hline
%      30 & \\ \hline
%      31 & \\ \hline
%      32 & \\ \hline
%      33 & \\ \hline
%      34 & \\ \hline
%      35 & \\ \hline
%      36 & \\ \hline
%      37 & \\ \hline
%    \end{tabular}
%    \egroup
%  \end{minipage}
%\end{center}
\newpage

\genfreetitle{2}{FlUID MECHANICS}{4}

\genfreedirections

% THIS TAKEN FROM THE 2001 AP PHYSICS B EXAM FREE-RESPPONSE QUESTION #5
\begin{center}
  \pic{.5}{submerged-block}\\
  \underline{Note:} Figure not drawn to scale.
\end{center}
\begin{enumerate}
\item A large rectangular raft (density \SI{650}{\kilo\gram\per\metre\cubed})
  is floating on a lake. The surface area of the top of the raft is
  \SI{8.2}{\metre\squared} and its volume is \SI{1.80}{\metre\cubed}. The
  density of the lake water is \SI{1000}{\kilo\gram\per\metre\cubed}.
  \begin{enumerate}
  \item Calculate the height $h$ of the portion of the raft that is above the
    surrounding water.
  \item Calculate the magnitude of the buoyant force on the raft and state its
    direction.
  \item If the average mass of a person is \SI{75}{\kilo\gram}, calculate the
    maximum number of people that can be on the raft without the top of the
    raft sinking below the surface of the water. (Assume that the people are
    evenly distributed on the raft.)
  \end{enumerate}
  \newpage
  
  % THIS IS TAKEN FROM THE 2012 AP PHYSICS B EXAM FREE-RESPPONSE QUESTION #3
  \cpic{.28}{oils}
  \item A glass U-tube with a uniform diameter of 0.850 cm is used to determine
  the density of an oil. As shown in the figure above, a 24.5 cm column of
  water balances a 27.2 cm column of the oil so that interfaces $A$ and $B$ of
  the mercury with the other liquids are at the same height. The density of
  water is \SI{1.00e3}{\kilo\gram\per\metre\cubed}.
  \begin{enumerate}
  \item Calculate the density of the oil.
    \vspace{2in}
  \item Calculate the absolute pressure at $B$, the interface between the water
    and the mercury.
    \vspace{2in}
  \end{enumerate}
  \cpic{.22}{u-tube}
  A new tube, identical to the U-tube except for a cone shape on the left, as
  shown above, is filled with the same volume of mercury that was in the
  U-tube. The mercury is at the same height on both sides of the new tube as it
  was in the U-tube, as shown by the dashed line. The same volumes of oil and
  water that were in the U-tube are now poured into the new tube, on the left
  and right respectively.
  \begin{enumerate}[resume]
  \item Indicate the new position of $B$ relative to $A$.
    
    \vspace{.1in}
    \underline{\hspace{.3in}} Above $A$\hspace{.2in}
    \underline{\hspace{.3in}} Below $A$\hspace{.2in}
    \underline{\hspace{.3in}} At the same height as $A$
    
    \vspace{.1in}Justify your answer.\vspace{1in}
      
  \item A small piece of wood with density less than that of the oil is placed
    so that it floats in the left side of the tube. Indicate whether the
    pressure at the bottom of the tube increases, decreases, or remains the
    same.

    \vspace{.1in}
    \underline{\hspace{.3in}} Increases\hspace{.2in}
    \underline{\hspace{.3in}} Decreases\hspace{.2in}
    \underline{\hspace{.3in}} Remains the same
  \end{enumerate}
  \newpage
  
  % THIS IS TAKEN FROM THE 2014 AP PHYSICS B EXAM FREE-RESPONSE QUESTION #2
  \begin{center}
    \begin{tikzpicture}[scale=1.5]
      \fill[gray!50](0,0) rectangle(2,2);
      \draw[very thick](0,2.5)--(0,0)--(2,0)--(2,2.5);
      \draw[very thick](0,2)--(2,2);
      \draw[very thick](1,0)--(1,1);
      \draw[very thick,fill=white](.8,1) rectangle(1.2,1.4);
    \end{tikzpicture}
  \end{center}
\item A cube of mass $m$ and side length $L$ is completely submerged in a tank
  of water and is attached to the bottom of the tank by a string, as shown in
  the figure above. The tension in the string is 0.25 times the weight of the
  cube. The density of water is \SI{1000}{\kilo\gram\per\metre\cubed}.
  \begin{enumerate}
  \item On the dot below that represents the cube, draw and label the forces
    (not components) that act on the cube while it is attached to the string.
    Each force must be represented by a distinct arrow starting on, and pointing
    away from, the dot.
    \begin{center}
      \vspace{.5in}
      \begin{tikzpicture}
        \fill(0,0) circle(.3);
      \end{tikzpicture}
      \vspace{.5in}
    \end{center}
  \item Calculate the density of the cube.
    \vspace{1.5in}

  \item The string is now cut. Calculate the magnitude of the
    acceleration of the cube immediately after the string is cut. If you need
    to draw anything other than what you have shown in part (a) to assist in
    your solution, use the space below. Do NOT add anything to the figure in
    part (a).
    \vspace{1.5in}
    
  \item Indicate whether the magnitude of the buoyant force on the cube
    increases, decreases, or remains the same while the cube is rising, but
    before it reaches the surface.

    \vspace{.1in}
    \underline{\hspace{.3in}} Increases\hspace{.2in}
    \underline{\hspace{.3in}} Decreases\hspace{.2in}
    \underline{\hspace{.3in}} Remains the same

    \vspace{.1in}Justify your answer.
  \end{enumerate}
  \newpage

  \cpic{.38}{fr-q2}
  
\item A \SI{1.}{\centi\metre} radius hose with a \SI{.50}{\centi\metre} radius
  exit nozzle is being used to fill a \SI{1000}{\milli\litre} beaker with oil\\
  (\SI{1000}{\milli\litre}=\SI{.0010}{\metre\cubed}). The velocity of the oil
  in the hose is $v=\SI{0.40}{\metre\per\second}$ as shown in the figure. The
  density of the oil is \SI{960}{\kilo\gram\per\metre\cubed}, and the
  atmospheric pressure is \SI{1.01e5}\pascal.
  
  \begin{enumerate}
  \item The nozzle attached to the end of the hose has a smaller radius than
    the hose. If the nozzle is removed from the hose, will the beaker be filled
    faster? Justify your answer with conservation laws.
    \vspace{\stretch1}
    
  \item Calculate the exit velocity of the oil from the nozzle.
    \vspace{\stretch1}
    
  \item How long will it take to fill the beaker?
    \vspace{\stretch1}
    
  \item Point A is shown in the figure. How does the pressure in the fluid at
    point A compare to the pressure in the fluid at the exit nozzle? Justify
    your claim.
    \vspace{\stretch1}
    
  \item The hose is now used to fill a \SI{200}{\milli\litre} graduated
    cylinder with oil to the same height as the height of the oil in the
    \SI{1000}{\milli\litre} beaker. Compare the net force from the oil on the
    bottom of the \SI{200}{\milli\litre} cylinder and the
    \SI{1000}{\milli\litre}. Explain your answer.
    \vspace{\stretch1}
  \end{enumerate}
  \newpage
  
\item A cube of lead with a side dimension of \SI{5.}{\centi\metre} is slowly
  lowered into the beaker of oil by a thin string attached to a spring scale at
  a constant rate, as shown in the figure. The density of lead is
  \SI{11300}{\kilo\gram\per\metre\cubed}.
  \cpic{.3}{fr-2a}
  
  \begin{enumerate}
  \item What will be the spring scale reading in newtons when the lead has
    been submerged to location 2?
    \vspace{\stretch1}
    
  \item Does the spring scale reading increase, decrease, or stay the same
    when the cube is lowered from location 2 to location 3? Justify your
    answer by referencing the pressure of the fluid on the lead cube.
    \vspace{\stretch1}
    
  \item The lead cube is lowered from above the oil's surface (location 1) to
    a spot just below the surface (location 2) until the cube is just above
    the bottom of the beaker (location 3). Describe any changes in pressure
    on the bottom of the beaker during this process. Explain your answer.
    \vspace{\stretch1}
  \end{enumerate}
\end{enumerate}
\end{document}
